### 2021-01-22 SpatialDesignVModel.Rmd
initial outline commit

---
title: 'Spatial Analysis: Design-Based vs Model-Based'
author: "Michael Dumelle and Matthew Higham"
output: 
  html_document:
    theme: flatly
    number_sections: true
    highlighted: default 
    toc: yes
    toc_float:
      collapsed: no
      smooth_scroll: no
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  collapse = TRUE,
  comment = "#>", 
  warning = FALSE, 
  message = FALSE
)
```

Insert Overview

# Outline

Here is an outline

## Introduction

Few sentences about the goal of the introduction

## Literature Review

Few sentences about the goal of the literature review

## Numerical Analysis

Few sentences about the goal of the data analysis

### Simulation Based

Few sentences about the goal of the simulations

### Data Based

Few sentences about the goal of the data analysis

## Discussion

Few sentences about the most important takeaways