### 2021-01-25 draft.Rnw
initial tex commit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document Start
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% this is where we read in all of the preamble material (packages, etc. - stuff that comes before the document begins)
\input{preamble.tex}

% starting the document
\begin{document}

% printing the title (which is in the preamble)
\maketitle

% adding line numbers
\linenumbers

% page break for abstract
\clearpage

% inserting abstract
% \input{abstract.tex}

% page break for start of manuscript
\clearpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Overview
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{\centering {Overview}}

\begin{itemize}
  \item Design-Based Overview
    \begin{itemize}
      \item A classic text overviewing design-based approaches \citep{Sarndal2003model}
      \item A classic text overviewing design-based approaches \citep{Lohr2009sampling}
    \end{itemize}
  \item Model-Based Overview
    \begin{itemize}
      \item A classic text overviewing model-based approaches to analyzing spatial data \citep{Cressie2015statistics}
      \item A classic text overviewing model-based approaches to analyzing spatial data \citep{Schabenberger2017statistical}
    \end{itemize}
  \item Model-based and Design-based Comparisons
      \begin{itemize}
        \item A comparison of design-based and model-based approaches \citep{Hansen1983evaluation}
        \item A comparison of design-based and model-based approaches \citep{Brus1997random}
        \item A comparison of design-based and model-based approaches for spatial data \citep{VerHoef2002sampling}
        \item A comparison of design-based and model-based approaches for spatial data \citep{Cooper2006sampling}
        \item A comparison of design-based and model-based approaches\citep{sterba2009alternative}
        \item A comparison of design-based and model-based approaches for spatial data \citep{Brus2020statistical}
        \item A Bayesian approach to design-based approaches for spatial data \citep{Chan2020bayesian}
      \end{itemize}
  \item Spatially Balanced Design and Analysis
      \begin{itemize}
        \item GRTS sampling \citep{StevensOlsen2004GRTS}
        \item Local neighborhood variance estimation \citep{StevensOlsen2003VarianceEstimation}
      \end{itemize}
  \item Block Kriging 
      \begin{itemize}
        \item Finite Population Block Kriging (with an overview of Block Kriging) \citep{VerHoef2008spatial}
      \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Outline
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{\centering OUTLINE}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Manuscript Start
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{\centering INTRODUCTION}

%----------------------
% Background
%----------------------
\section{\centering BACKGROUND}

%----------------------
% Numerical Analysis
%----------------------
\section{\centering NUMERICAL ANALYSIS}

\subsection{Simulation-Based}

\subsection{Data-Based}

\subsection{Software}

%----------------------
% Discussion
%----------------------
\section{\centering DISCUSSION}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Document End
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% page break for bibliography
\clearpage

% bibliography style
\bibliographystyle{apalike}

% bibliography
\bibliography{draft}

% ending the document
\end{document}