### 2021-06-09 inst/SpatialDVM_Manuscript/SpatialDVM_Manuscript.tex
adjusted grts section and added simulation one table

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{lineno} % add
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing

% Pandoc header

\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}}

\begin{document}
\begin{frontmatter}

  \title{A comparison of design-based and model-based approaches for spatial
data.}
    \author[USEPA]{Michael Dumelle\corref{1}}
   \ead{Dumelle.Michael@epa.gov} 
    \author[STLAW]{Matthew Higham\corref{1}}
   \ead{mhigham@stlaw.edu} 
    \author[OSU]{Lisa Madsen}

    \author[USEPA]{Anthony R. Olsen}

    \author[NOAA]{Jay M. Ver Hoef}

      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
Corvallis, Oregon, 97333}
    \address[STLAW]{Saint Lawrence University Department of Math, Computer Science, and
Statistics, 23 Romoda Drive, Canton, New York, 13617}
    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
Corvallis, Oregon, 97331}
    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
      \cortext[1]{Corresponding Author}

  \begin{abstract}
  This is the abstract.

  It consists of two paragraphs.
  \end{abstract}

 \end{frontmatter}

\emph{Text based on elsarticle sample manuscript, see
\url{http://www.elsevier.com/author-schemas/latex-instructions\#elsarticle}}

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

finish spatial introduction revise section 2 reread brus and wang write
down potential data sets

Please leave comments in your
\textcolor{blue}{c}\textcolor{magenta}{o}\textcolor{green}{l}\textcolor{red}{o}\textcolor{cyan}{r}:
\textcolor{blue}{Michael}, \textcolor{magenta}{Matt},
\textcolor{green}{Lisa}, \textcolor{red}{Tony}, \textcolor{cyan}{Jay}.

There are two general approaches for using data to make statistical
inferences about a population: design-based approaches and model-based
approaches. When data cannot be obtained for all units in a population
(known as population units), data on a subset of the population units is
collected in a sample. In the design-based approach, inferences about
the underlying population are informed from a probabilistic process in
which population units are selected to be in the sample. Alternatively,
in the model-based approach, inferences are made from specific
assumptions made about the underlying process that generated the data.
Each paradigm has a deep historical context (Sterba, 2009) and its own
set of general advantages (Hansen et al., 1983).

Though the design-based and model-based approaches apply to statistical
inference in a broad sense, we focus on comparing these approaches for
spatial data. We define spatial data as variables measured at specific
geographic locations. De Gruijter and Ter Braak (1990) give an early
comparison of design-based and model-based approaches for spatial data,
quashing the belief that design-based approaches could not be used for
spatially correlated data. Thereafter, several comparisons between
design-based and model-based for spatial data have been considered, but
they tend to compare design-based approaches that ignore spatial
location in sampling to model-based approaches (Brus and De Gruijter,
1997; Ver Hoef, 2002, 2008). More recent overviews include Brus (2020)
and Wang et al. (2012), but no numerical comparison has been made
between design-based approaches that incorporate spatial location in
sampling and model-based approaches.

The rest of this paper is organized as follows. In Section
\ref{sec:background}, we compare sampling and estimation procedures
between the design-based approach and the model-based approach. In
Section \ref{sec:numstudy}, we use simulated and real data to study the
properties of parameter estimates from both approaches. And in Section
\ref{sec:discussion}, we end with a discussion and provide directions
for future research.

\hypertarget{sec:background}{%
\section{Background}\label{sec:background}}

The design-based and model-based approaches incorporate randomness in
fundamentally different ways. In this section, we describe the role of
randomness and its effects on subsequent inferences. We then discuss
specific inference methods for the design-based and model-based
approaches for spatial data.

\hypertarget{comparing-design-based-vs.-model-based}{%
\subsection{Comparing Design-Based
vs.~Model-Based}\label{comparing-design-based-vs.-model-based}}

\hypertarget{design-based-philosphy}{%
\subsubsection{Design-Based Philosphy}\label{design-based-philosphy}}

The design-based approach assumes the data are fixed. Randomness is
incorporated in the selection of population units according to a
sampling design. A sampling design assigns a positive probability of
inclusion in the sample (inclusion probability) to each population unit.
Some examples of commonly used sampling designs include independent
random sampling (IRS), stratified random sampling, and cluster sampling.
The goal is to use the sampling design and the sampled data to estimate
population parameters like means and totals. These population parameters
are typically assumed to be fixed but unknown.

Treating the data as fixed and incorporating randomness through the
sampling design yields estimators having very few other assumptions.
Confidence intervals for these types of estimators are typically derived
using limiting arguments. Means and totals, for example, are
asymptotically normally distributed by the Central Limit Theorem.
Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
design-based approach.

\hypertarget{model-based-philosphy}{%
\subsubsection{Model-Based Philosphy}\label{model-based-philosphy}}

The model-based approach assumes the data are a random realization of a
process. Randomness is often incorporated through distributional
assumptions on the data-generating process. Instead of estimating fixed
but unknown parameters (as in the design-based approach), the goal of
model-based inference in the spatial context is often \emph{prediction}
of an unknown quantity. For example, suppose the realized mean of all
population units is the quantity of interest. Instead of
\emph{estimating} a fixed unknown mean, we are \emph{predicting} the
value of the mean, a random variable. We know that if we sampled all
population units, we would have an exact prediction for the mean of our
one realized process, without any uncertainty. But the true mean of the
spatial process that generated our realized data is still not known.
When predicting the realized mean, we typically are not interested in
the underlying process' true mean.

Assuming the data is a realization of a specific data-generating process
yields predictors that are linked to distributional assumptions. These
distributional assumptions are used to derive prediction intervals. The
distributional assumptions allow the prediction intervals can be more
precise. Cressie (1993) and Schabenberger and Gotway (2017) provide
reviews of model-based approaches for spatial data.

Several comparisons between the model-based paradigm and the
design-based paradigm have been made in many different contexts. Sterba
(2009) give some history of the two paradigms as well as some
applications in psychological contexts. Cooper (2006) review the two
approaches in an ecological context before introducing a
``model-assisted'' variance estimator that combines aspects from each
approach. We note that while there has been substantial research and
development into estimators that use both design and model-based
principles (see e.g.~Cicchitelli and Montanari (2012), Chan-Golston et
al. (2020) for a Bayesian approach, CITE some more of these), our goal
is not to expand upon or improve these methods.

Figure 1a. Brus (2020): Data is fixed. In a finite population example,
show a 3d surface that can be generated by anything. If we repeatedly
sample the surface, then 95\% of all 95\% CIs will contain the true
mean, which never changes.

\includegraphics[width=0.85\linewidth]{SpatialDVM_Manuscript_files/figure-latex/unnamed-chunk-2-1}

Figure 1b. Spatial process is fixed. In a finite population example,
show 10 3d surfaces that are generated from some model. If we repeatedly
generate the surface and obtain a sample, then 95\% of all 95\% PIs will
contain the realized means. The realized mean changes from surface to
surface and it's not necessarily the case that 95\% of all 95\% PIs will
contain the true, underlying mean.

\hypertarget{spatially-balanced-design-and-analysis}{%
\subsection{Spatially Balanced Design and
Analysis}\label{spatially-balanced-design-and-analysis}}

Sampling designs can incorporate spatial locations to obtain samples
that are spatially balanced with respect to the population (Stevens Jr
and Olsen, 2004). A sample is spatially with respect to the population
if the sampled population units are a miniature of the population units.
A sample is a miniature of the population if the distribution of the
sampled population units mirrors the density of all population units.
Spatial balance with respect to the population is is different than
spatial balance with respect to geography. A sample that is spatially
balanced with respect to geography is spread out in some type of
equidistant manner over geographical space and are not meant to be
miniatures of the population. When we refer to spatial balance
henceforth, we mean spatial balance with respect to the population.

Spatially balanced samples are useful because they tend to yield
estimates that have lower variance than estimates constructed from
sampling designs lacking spatial balance (Barabesi and Franceschi, 2011;
Benedetti et al., 2017; Grafström and Lundström, 2013; Robertson et al.,
2013; Stevens Jr and Olsen, 2004; Wang et al., 2013). To quantify
spatial balance, Stevens Jr and Olsen (2004) proposed loss functions
based on Voroni polygons. The first spatiallly balanced sampling
algorithm that saw widespread use was the Generalized Random
Tessellation Stratified (Stevens Jr and Olsen, 2004). Since GRTS was
developed, several other spatially balanced sampling algorithms have
emerged, including the Local Pivotal Method (Grafström et al., 2012;
Grafström and Matei, 2018), Spatially Correlated Poisson Sampling
(Grafström, 2012), Balanced Acceptance Sampling (Robertson et al.,
2013), Within-Sample-Distance (Benedetti and Piersimoni, 2017), and
Halton Iterative Partitioning (Robertson et al., 2018) algorithms. We
focus on the Generalized Random Tessellation Stratified (GRTS) algorithm
to select spatially balanced sampling because the algorithm has several
attractive properties detailed by Stevens Jr and Olsen (2004) and
Dumelle et al. (2021).

The GRTS algorithm is used to sample from finite and infinite
populations and works by utilizing a mapping between two-dimensional and
one-dimensional space. The population units in two-dimensional space are
divided into cells using a hierarchical index. Population units are then
mapped to a one-dimensional line via the hierarchical indexing. The line
length of each population unit equals its inclusion probability. A
systematic sample is conducted on the line and these samples are linked
to a population unit in two-dimensional space, which results in the
desired sample size. Stevens Jr and Olsen (2004) provide and Dumelle et
al. (2021) provide further details. The GRTS algorithm is available in
the \textbf{\textsf{R}} package \texttt{spsurvey} (Dumelle et al.,
2021).

After collecting a sample, the data are used to estimate population
parameters. The Horvitz-Thompson estimator (Horvitz and Thompson, 1952)
and its continuous analog (Cordy, 1993) yield unbiased estimates of
population means and totals. If \(\tau\) is a population total, then the
Horvitz-Thompson estimator of \(\tau\) (denoted by \(\hat{\tau}_{ht}\)),
is given by \begin{align}\label{eq:ht}
  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
\end{align} where \(Z_i\) and \(pi_i\) are the observed value and
inclusion probability of the \(i\)th population unit selected in the
sample. Horvitz and Thompson (1952) and Sen (1953) provide variance
estimators for \(\hat{\tau}_{ht}\), but they have two drawbacks. First,
they rely on calculating \(\pi_{ij}\), the probability that population
unit \(i\) and population unit \(j\) are included in the sample, which
can be very difficult to calculate. Second, they ignore the spatial
locations of the population units. To address these drawbacks, Stevens
Jr and Olsen (2003) proposed a local neighborhood variance estimator.
The local neighborhood variance estimator does not rely on \(\pi_{ij}\),
and it incorporates spatial locations by assigning higher weights to
nearby observations. Stevens Jr and Olsen (2003) show this variance
estimators tends to reduce the variability associated with estimating
\(\tau\). This yields confidence intervals for \(\tau\) that are
narrower than confidence intervals constructed from variance estimators
ignoring spatial locations.

\hypertarget{finite-population-block-kriging}{%
\subsection{Finite Population Block
Kriging}\label{finite-population-block-kriging}}

We only use FPBK in this paper in order to focus more on comparing the
design-based and model-based approaches. However, k-nearest-neighbors
(Fix and Hodges, 1951; Ver Hoef and Temesgen, 2013), random forest
(Breiman, 2001), Bayesian models (Chan-Golston et al., 2020), among
others, can also be used to obtain predictions for a mean or total from
spatially correlated responses in a finite population setting.

Finite Population Block Kriging (FPBK) is an alternative to
samipling-based methods (Ver Hoef, 2008). FPBK expands the
geostatistical kriging framework to the finite population setting.
Instead of basing inference off of a specific sampling design, we assume
the data were generated by a spatial process with parameters that can be
estimated using the framework of a model.

Ver Hoef (2008) gives details on the theory of FPBK, but some of the
basic principles are summarized below. For a response variable
\(\mathbf{z}\) that can be measured on a finite number of \(N\) sites,
we want to predict some linear function of the response variable,
\(\tau(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
\(\mathbf{b}\) is a vector of weights. For example, if we want to
predict the total abundance across all sites, then we would use a vector
of 1's for the weights.

Typically, however, we only have a sample of the \(N\) sites. Denoting
quantities that are part of the sampled sites with a subscript \emph{s}
and quantities that are part of the unsampled sites with a subscript
\emph{u},

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
design matrices for the sampled and unsampled sites, respectively, and
\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
sampled and unsampled sites. Denoting
\(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), we assume
that \(E(\bm{\delta})\) = \(\mathbf{0}\).

We also typically assume that there is spatial correlation in
\(\bm{\delta}\), which can be modeled using a covariance function. Many
common choices for this function assume that spatial covariance
decreases with increasing Euclidean distance between sites. The primary
function used throughout the simulations and applications of this
manuscript is the Exponential covariance function: the \(i,j^{th}\)
entry for \(\mathop{\mathrm{{var}}}(\bm{\delta})\) is \mbox{}
\begin{align}\label{equation:expcov}
\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = \theta_1\exp(-3h_{i,j}/\theta_2) + \theta_3\mathbbm{1}\{\mathbf{h}_{i,j} = 0\}, 
\end{align} where \(h_{i,j}\) is the distance between sites \(i\) and
\(j\), and \(\bm{\theta}\) is a vector of spatial covariance parameters
of the partial sill \(\theta_1\), the range \(\theta_2\), and the nugget
\(\theta_3\). However, any spatial covariance function could be used in
the place of the Exponential, including functions that allow for
anisotropy {[}pg. 80 - 93{]}(Chiles and Delfiner, 1999).

With the above model formulation, the Best Linear Unbiased Predictor
(BLUP) for \(\tau(\mathbf{b}'\mathbf{z})\)
\(\tau(\mathbf{b}'\mathbf{z})\) is vague\} and its prediction variance
can be computed. While details of the derivation are in (Ver Hoef,
2008), we note here that the predictor and its variance are both
moment-based. Neither require a particular distribution for
\(\mathbf{z}\).

\hypertarget{sec:numstudy}{%
\section{Numerical Study}\label{sec:numstudy}}

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrrr}
  \hline
Approach & Bias & RMSE & MedAE & Coverage & PClose & MedIL \\ 
  \hline
Design & 0.0003 & 0.0353 & 0.0251 & 0.9461 & 0.4889 & 0.1362 \\ 
  Model & -0.0001 & 0.0362 & 0.0253 & 0.9480 & 0.5111 & 0.1430 \\ 
   \hline
\end{tabular}
\caption{Approach, mean bias (Bias), root-mean-squared error (RMSE), median absolute error (MedAE), 95 percent interval coverage (Coverage), proportion of times the approach was closer to the true value (PClose), and median interval length (MedIL)} 
\label{tab:sim_one}
\end{table}

\textbf{Base Simulations}

\begin{itemize}
\item
  both good: correctly specified model with high correlation
\item
  break model: non-gaussian errors
\item
  break design: small area estimation
\item
  both good?: misspecified covariance model with high correlation
\item
  break both? non-gaussian areas with smaller sample size
\item
  change n or sampling fraction
\item
  model-based: how should sample be drawn? should locations be fixed?
\end{itemize}

\hypertarget{software}{%
\subsection{Software}\label{software}}

FPBK can be readily performed in \texttt{R} with the \texttt{sptotal}
package (Higham et al., 2020). We use \texttt{sptotal} for both the
simulation analysis and the application, estimating parameters with
Restricted Maximum Likelihood (REML).

\hypertarget{applied-example}{%
\subsection{Applied Example}\label{applied-example}}

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-barabesi2011sampling}{}%
Barabesi, L., Franceschi, S., 2011. Sampling properties of spatial total
estimators under tessellation stratified designs. Environmetrics 22,
271--278.

\leavevmode\hypertarget{ref-benedetti2017spatially}{}%
Benedetti, R., Piersimoni, F., 2017. A spatially balanced design with
probability function proportional to the within sample distance.
Biometrical Journal 59, 1067--1084.

\leavevmode\hypertarget{ref-benedetti2017spatiallyreview}{}%
Benedetti, R., Piersimoni, F., Postiglione, P., 2017. Spatially balanced
sampling: A review and a reappraisal. International Statistical Review
85, 439--454.

\leavevmode\hypertarget{ref-breiman2001random}{}%
Breiman, L., 2001. Random forests. Machine learning 45, 5--32.

\leavevmode\hypertarget{ref-brus1997random}{}%
Brus, D., De Gruijter, J., 1997. Random sampling or geostatistical
modelling? Choosing between design-based and model-based sampling
strategies for soil (with discussion). Geoderma 80, 1--44.

\leavevmode\hypertarget{ref-brus2020statistical}{}%
Brus, D.J., 2020. Statistical approaches for spatial sample survey:
Persistent misconceptions and new developments. European Journal of Soil
Science.

\leavevmode\hypertarget{ref-chan2020bayesian}{}%
Chan-Golston, A.M., Banerjee, S., Handcock, M.S., 2020. Bayesian
inference for finite populations under spatial process settings.
Environmetrics 31, e2606.

\leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
Chiles, J.-P., Delfiner, P., 1999. Geostatistics: Modeling Spatial
Uncertainty. John Wiley \& Sons, New York.

\leavevmode\hypertarget{ref-cicchitelli2012model}{}%
Cicchitelli, G., Montanari, G.E., 2012. Model-assisted estimation of a
spatial population mean. International Statistical Review 80, 111--126.

\leavevmode\hypertarget{ref-cooper2006sampling}{}%
Cooper, C., 2006. Sampling and variance estimation on continuous
domains. Environmetrics: The official journal of the International
Environmetrics Society 17, 539--553.

\leavevmode\hypertarget{ref-cordy1993extension}{}%
Cordy, C.B., 1993. An extension of the horvitz---thompson theorem to
point sampling from a continuous universe. Statistics \& Probability
Letters 18, 353--362.

\leavevmode\hypertarget{ref-cressie1993statistics}{}%
Cressie, N., 1993. Statistics for spatial data. John Wiley \& Sons.

\leavevmode\hypertarget{ref-de1990model}{}%
De Gruijter, J., Ter Braak, C., 1990. Model-free estimation from spatial
samples: A reappraisal of classical sampling theory. Mathematical
geology 22, 407--415.

\leavevmode\hypertarget{ref-dumelle2021spsurvey}{}%
Dumelle, M., Olsen, A.R., Kincaid, T., Weber, M., 2021. Selecting and
analyzing spatial probability samples in r using spsurvey. Manuscript
Submitted for Publication.

\leavevmode\hypertarget{ref-fix1951discriminatory}{}%
Fix, E., Hodges, J.L., 1951. Discriminatory analysis, nonparametric
discrimination: Consistency properties. USAF School of Aviation
Medicine.

\leavevmode\hypertarget{ref-grafstrom2012spatiallypoisson}{}%
Grafström, A., 2012. Spatially correlated poisson sampling. Journal of
Statistical Planning and Inference 142, 139--147.

\leavevmode\hypertarget{ref-grafstrom2013well}{}%
Grafström, A., Lundström, N.L., 2013. Why well spread probability
samples are balanced. Open Journal of Statistics 3, 36--41.

\leavevmode\hypertarget{ref-grafstrom2012spatially}{}%
Grafström, A., Lundström, N.L., Schelin, L., 2012. Spatially balanced
sampling through the pivotal method. Biometrics 68, 514--520.

\leavevmode\hypertarget{ref-grafstrom2018spatially}{}%
Grafström, A., Matei, A., 2018. Spatially balanced sampling of
continuous populations. Scandinavian Journal of Statistics 45, 792--805.

\leavevmode\hypertarget{ref-hansen1983evaluation}{}%
Hansen, M.H., Madow, W.G., Tepping, B.J., 1983. An evaluation of
model-dependent and probability-sampling inferences in sample surveys.
Journal of the American Statistical Association 78, 776--793.

\leavevmode\hypertarget{ref-higham2020sptotal}{}%
Higham, M., Ver Hoef, J., Bryce, F., 2020. Sptotal: Predicting totals
and weighted sums from spatial data.

\leavevmode\hypertarget{ref-horvitz1952generalization}{}%
Horvitz, D.G., Thompson, D.J., 1952. A generalization of sampling
without replacement from a finite universe. Journal of the American
statistical Association 47, 663--685.

\leavevmode\hypertarget{ref-lohr2009sampling}{}%
Lohr, S.L., 2009. Sampling: Design and analysis. Nelson Education.

\leavevmode\hypertarget{ref-robertson2013bas}{}%
Robertson, B., Brown, J., McDonald, T., Jaksons, P., 2013. BAS: Balanced
acceptance sampling of natural resources. Biometrics 69, 776--784.

\leavevmode\hypertarget{ref-robertson2018halton}{}%
Robertson, B., McDonald, T., Price, C., Brown, J., 2018. Halton
iterative partitioning: Spatially balanced sampling via partitioning.
Environmental and Ecological Statistics 25, 305--323.

\leavevmode\hypertarget{ref-sarndal2003model}{}%
Särndal, C.-E., Swensson, B., Wretman, J., 2003. Model assisted survey
sampling. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-schabenberger2017statistical}{}%
Schabenberger, O., Gotway, C.A., 2017. Statistical methods for spatial
data analysis. CRC press.

\leavevmode\hypertarget{ref-sen1953estimate}{}%
Sen, A.R., 1953. On the estimate of the variance in sampling with
varying probabilities. Journal of the Indian Society of Agricultural
Statistics 5, 127.

\leavevmode\hypertarget{ref-sterba2009alternative}{}%
Sterba, S.K., 2009. Alternative model-based and design-based frameworks
for inference from samples to populations: From polarization to
integration. Multivariate behavioral research 44, 711--740.

\leavevmode\hypertarget{ref-stevens2003variance}{}%
Stevens Jr, D.L., Olsen, A.R., 2003. Variance estimation for spatially
balanced samples of environmental resources. Environmetrics 14,
593--610.

\leavevmode\hypertarget{ref-stevens2004spatially}{}%
Stevens Jr, D.L., Olsen, A.R., 2004. Spatially balanced sampling of
natural resources. Journal of the american Statistical association 99,
262--278.

\leavevmode\hypertarget{ref-verhoef2002sampling}{}%
Ver Hoef, J., 2002. Sampling and geostatistics for spatial data.
Ecoscience 9, 152--161.

\leavevmode\hypertarget{ref-verhoef2008spatial}{}%
Ver Hoef, J.M., 2008. Spatial methods for plot-based sampling of
wildlife populations. Environmental and Ecological Statistics 15, 3--13.

\leavevmode\hypertarget{ref-ver2013comparison}{}%
Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
model to nearest neighbor (k-nn) methods for forestry applications. PloS
one 8, e59129.

\leavevmode\hypertarget{ref-wang2013design}{}%
Wang, J.-F., Jiang, C.-S., Hu, M.-G., Cao, Z.-D., Guo, Y.-S., Li, L.-F.,
Liu, T.-J., Meng, B., 2013. Design-based spatial sampling: Theory and
implementation. Environmental modelling \& software 40, 280--288.

\leavevmode\hypertarget{ref-wang2012review}{}%
Wang, J.-F., Stein, A., Gao, B.-B., Ge, Y., 2012. A review of spatial
sampling. Spatial Statistics 2, 1--14.

\end{document}