### 2021-06-25 inst/SpatialDVM_Manuscript/notes_to_do.Rmd
added to-do list with notes from June 24 meeting.

---
title: "To Do List"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction and Background:

* consider re-wording the model-based use of the word "parameter" to make the parameter generating the process distinct from using "parameter" as, for example, the realized mean (which we treat as fixed for one particular realization in the finite population setting).

Simulations:

* put in SRS as base-line sampling setting.
* add FPBK as traditional SRS analysis as methods to analyze SRS sample.
* add FPBK and traditional SRS analysis as additional methods to analyze GRTS sample.
* add GRTS with 10% nugget sites as additional sampling setting? If so, add FPBK to analyze this sample.