### 2021-10-12 inst/manuscript/notes_Madsen.Rmd
updates to section 1 and 2

# Please Leave Your Notes Here

__Lisa M__: Add a paragraph describing the contribution of the manuscript (before the "the rest of the manuscript is organize as follows etc." in the introduction)