### 2021-10-12 inst/manuscript/notes_Madsen.Rmd
section 2 updates

# Please Leave Your Notes Here

__Lisa M__: Add a paragraph describing the contribution of the manuscript (before the "the rest of the manuscript is organize as follows etc." in the introduction)

__Lisa M.__: Need a more precise definition of "miniature" in this context, and need an example.

__Lisa M.__: List major attractive properties, and detailed by @stevens2004spatially and @dumelle2021spsurvey.

__Lisa M.__ : Include formulas. Perhaps, but, these are very heavy in notation and matrix algebra. We might consider, however, adding the formulas to an Appendix.