### 2021-10-13 inst/manuscript/notes_Higham.Rmd
section 3 and 4 edits

Notes for Matthew Higham

- for the model-based approach, the more skewed the population is, the higher the sample size needed to satisfy CLT for predicting a mean. The derivation of the BLUP is entirely moment-based (no distribution assumed) but we still need to assume a distribution to estimate spatial parameters and to generate bounds of a prediction interval.
- many confidence intervals generated for design-based approaches also rely on the CLT and the normal distribution to generate the interval. Again, for highly skewed data with a small sample size, this assumption is violated even though all of the assumptions for generating the estimator are valid.

-Although all lakes in the survey were measured in 2012, there may not always be enough time or money to do so.

- (mike: may get in trouble for this one) If it is acceptable to have an interval for mean mercury concentration of about 25 ng / g and if we ignore the other variables that the EPA collects information on in these NLA surveys, then the EPA could consider sampling only 100 lakes in the future to save time and money.