### 2021-12-22 inst/manuscript/supporting.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Methods in Ecology and Evolution} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{graphicx}
\usepackage{booktabs} % book-quality tables
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supporting Information.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% Pandoc citation processing

% Pandoc header

\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \renewenvironment{abstract}{}{}

\begin{document}
\begin{frontmatter}

  \title{A comparison of design-based and model-based approaches for finite
population spatial data -- Supporting Information.}
    \author[USEPA]{Michael Dumelle\corref{1}}

    \author[STLAW]{Matt Higham}

    \author[OSU]{Lisa Madsen}

    \author[USEPA]{Anthony R. Olsen}

    \author[NOAA]{Jay M. Ver Hoef}

      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
Corvallis, Oregon, 97333}
    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
and Statistics, 23 Romoda Drive, Canton, New York, 13617}
    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
Corvallis, Oregon, 97331}
    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}

  \begin{abstract}

  \end{abstract}

 \end{frontmatter}

\hypertarget{sec:simtabs}{%
\section{Simulation Study Tables}\label{sec:simtabs}}

Recall the four sampling-analysis combinations from the simulation
study, summarized in Table \ref{tab:designanalysis}. Also recall the 36
parameter configurations from the simulation study, summarized in Table
\ref{tab:parmtab}.

\begin{table}[ht]
\centering
\begin{tabular}{r|ll}
  \hline
 & Design & Model \\ 
  \hline
IRS & IRS-Design & IRS-Model \\ 
  GRTS & GRTS-Design & GRTS-Model \\ 
   \hline
\end{tabular}
\caption{\label{tab:designanalysis} Sampling-analysis combinations in the simulation study. The rows give the two types of sampling designs and the columns give the two types of analyses.} 
\end{table}

\begin{table}[ht]
\centering
\begin{tabular}{r|lll}
   \hline
Sample Size (n) & 50 & 100 & 200 \\ 
  Layout & Random & Gridded & - \\ 
  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
  Response Type & Normal & Lognormal & - \\ 
   \hline
\end{tabular}
\caption{\label{tab:parmtab} Simulation scenario options. All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios. In each simualtion scenario, the total variance was 2.} 
\end{table}

Next we present tables summarizing mean estimation bias (design-based)
or mean prediction bias (model-based), mean standard errors,
root-mean-squared error (design-based) or root-mean-squared-prediction
error (model-based), and 95\% confidence interval coverage
(design-based) or 95\% prediction interval coverage (model-based) for
all 36 simulation scenarios.

\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0 & 50 & -0.0023 & 0.1968 & 0.1936 & 0.9440 \\ 
  IRS-Model & Gridded & Normal & 0 & 50 & -0.0029 & 0.1988 & 0.1917 & 0.9400 \\ 
  GRTS-Design & Gridded & Normal & 0 & 50 & -0.0011 & 0.1946 & 0.1703 & 0.9110 \\ 
  GRTS-Model & Gridded & Normal & 0 & 50 & -0.0006 & 0.1949 & 0.1875 & 0.9300 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 1.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0 & 50 & -0.0036 & 0.2007 & 0.1855 & 0.9080 \\ 
  IRS-Model & Gridded & Lognormal & 0 & 50 & -0.0094 & 0.2071 & 0.1833 & 0.9045 \\ 
  GRTS-Design & Gridded & Lognormal & 0 & 50 & -0.0007 & 0.1962 & 0.1629 & 0.8690 \\ 
  GRTS-Model & Gridded & Lognormal & 0 & 50 & -0.0022 & 0.1969 & 0.1787 & 0.8945 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 2.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0 & 50 & 0.0019 & 0.1861 & 0.1935 & 0.9505 \\ 
  IRS-Model & Random & Normal & 0 & 50 & 0.0014 & 0.1892 & 0.1912 & 0.9445 \\ 
  GRTS-Design & Random & Normal & 0 & 50 & -0.0040 & 0.1955 & 0.1709 & 0.9090 \\ 
  GRTS-Model & Random & Normal & 0 & 50 & -0.0040 & 0.1965 & 0.1879 & 0.9315 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 3.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0 & 50 & 0.0032 & 0.1843 & 0.1833 & 0.9205 \\ 
  IRS-Model & Random & Lognormal & 0 & 50 & -0.0039 & 0.1945 & 0.1808 & 0.9105 \\ 
  GRTS-Design & Random & Lognormal & 0 & 50 & -0.0056 & 0.1944 & 0.1634 & 0.8870 \\ 
  GRTS-Model & Random & Lognormal & 0 & 50 & -0.0078 & 0.1967 & 0.1789 & 0.9075 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 4.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.5 & 50 & 0.0025 & 0.1762 & 0.1748 & 0.9470 \\ 
  IRS-Model & Gridded & Normal & 0.5 & 50 & 0.0007 & 0.1655 & 0.1567 & 0.9305 \\ 
  GRTS-Design & Gridded & Normal & 0.5 & 50 & 0.0004 & 0.1524 & 0.1357 & 0.9115 \\ 
  GRTS-Model & Gridded & Normal & 0.5 & 50 & -0.0003 & 0.1499 & 0.1453 & 0.9320 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 5.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0004 & 0.1863 & 0.1627 & 0.9140 \\ 
  IRS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0076 & 0.1834 & 0.1495 & 0.9035 \\ 
  GRTS-Design & Gridded & Lognormal & 0.5 & 50 & 0.0016 & 0.1612 & 0.1303 & 0.8810 \\ 
  GRTS-Model & Gridded & Lognormal & 0.5 & 50 & -0.0017 & 0.1606 & 0.1385 & 0.8940 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 6.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.5 & 50 & 0.0037 & 0.1657 & 0.1737 & 0.9590 \\ 
  IRS-Model & Random & Normal & 0.5 & 50 & 0.0027 & 0.1546 & 0.1559 & 0.9495 \\ 
  GRTS-Design & Random & Normal & 0.5 & 50 & -0.0034 & 0.1511 & 0.1355 & 0.9170 \\ 
  GRTS-Model & Random & Normal & 0.5 & 50 & -0.0037 & 0.1504 & 0.1447 & 0.9305 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 7.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.5 & 50 & 0.0023 & 0.1637 & 0.1574 & 0.9245 \\ 
  IRS-Model & Random & Lognormal & 0.5 & 50 & -0.0055 & 0.1622 & 0.1443 & 0.9125 \\ 
  GRTS-Design & Random & Lognormal & 0.5 & 50 & -0.0057 & 0.1570 & 0.1287 & 0.9000 \\ 
  GRTS-Model & Random & Lognormal & 0.5 & 50 & -0.0079 & 0.1567 & 0.1362 & 0.9100 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 8.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.9 & 50 & 0.0053 & 0.1579 & 0.1570 & 0.9470 \\ 
  IRS-Model & Gridded & Normal & 0.9 & 50 & 0.0026 & 0.1165 & 0.1094 & 0.9315 \\ 
  GRTS-Design & Gridded & Normal & 0.9 & 50 & 0.0013 & 0.1074 & 0.0996 & 0.9220 \\ 
  GRTS-Model & Gridded & Normal & 0.9 & 50 & -0.0007 & 0.0949 & 0.0938 & 0.9430 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 9.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1731 & 0.1412 & 0.9220 \\ 
  IRS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0020 & 0.1325 & 0.1051 & 0.9135 \\ 
  GRTS-Design & Gridded & Lognormal & 0.9 & 50 & 0.0031 & 0.1183 & 0.0954 & 0.9065 \\ 
  GRTS-Model & Gridded & Lognormal & 0.9 & 50 & -0.0002 & 0.1090 & 0.0909 & 0.9120 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 10.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.9 & 50 & 0.0062 & 0.1520 & 0.1545 & 0.9525 \\ 
  IRS-Model & Random & Normal & 0.9 & 50 & 0.0037 & 0.1074 & 0.1078 & 0.9525 \\ 
  GRTS-Design & Random & Normal & 0.9 & 50 & -0.0029 & 0.1038 & 0.0984 & 0.9340 \\ 
  GRTS-Model & Random & Normal & 0.9 & 50 & -0.0027 & 0.0940 & 0.0921 & 0.9360 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 11.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.9 & 50 & 0.0053 & 0.1542 & 0.1351 & 0.9325 \\ 
  IRS-Model & Random & Lognormal & 0.9 & 50 & -0.0012 & 0.1160 & 0.0997 & 0.9115 \\ 
  GRTS-Design & Random & Lognormal & 0.9 & 50 & -0.0033 & 0.1150 & 0.0923 & 0.9160 \\ 
  GRTS-Model & Random & Lognormal & 0.9 & 50 & -0.0031 & 0.1014 & 0.0872 & 0.9235 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 12.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0 & 100 & -0.0029 & 0.1322 & 0.1333 & 0.9430 \\ 
  IRS-Model & Gridded & Normal & 0 & 100 & -0.0033 & 0.1332 & 0.1326 & 0.9410 \\ 
  GRTS-Design & Gridded & Normal & 0 & 100 & 0.0041 & 0.1302 & 0.1208 & 0.9245 \\ 
  GRTS-Model & Gridded & Normal & 0 & 100 & 0.0038 & 0.1304 & 0.1309 & 0.9420 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 13.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0 & 100 & -0.0052 & 0.1334 & 0.1309 & 0.9335 \\ 
  IRS-Model & Gridded & Lognormal & 0 & 100 & -0.0072 & 0.1350 & 0.1301 & 0.9300 \\ 
  GRTS-Design & Gridded & Lognormal & 0 & 100 & 0.0042 & 0.1303 & 0.1170 & 0.9035 \\ 
  GRTS-Model & Gridded & Lognormal & 0 & 100 & 0.0034 & 0.1307 & 0.1266 & 0.9265 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 14.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0 & 100 & -0.0066 & 0.1366 & 0.1329 & 0.9405 \\ 
  IRS-Model & Random & Normal & 0 & 100 & -0.0067 & 0.1370 & 0.1322 & 0.9385 \\ 
  GRTS-Design & Random & Normal & 0 & 100 & 0.0028 & 0.1364 & 0.1210 & 0.9180 \\ 
  GRTS-Model & Random & Normal & 0 & 100 & 0.0029 & 0.1363 & 0.1310 & 0.9345 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 15.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0 & 100 & -0.0051 & 0.1348 & 0.1298 & 0.9280 \\ 
  IRS-Model & Random & Lognormal & 0 & 100 & -0.0075 & 0.1372 & 0.1290 & 0.9230 \\ 
  GRTS-Design & Random & Lognormal & 0 & 100 & 0.0002 & 0.1375 & 0.1181 & 0.8840 \\ 
  GRTS-Model & Random & Lognormal & 0 & 100 & -0.0003 & 0.1377 & 0.1275 & 0.9090 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 16.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.5 & 100 & -0.0016 & 0.1177 & 0.1201 & 0.9540 \\ 
  IRS-Model & Gridded & Normal & 0.5 & 100 & -0.0023 & 0.1072 & 0.1057 & 0.9470 \\ 
  GRTS-Design & Gridded & Normal & 0.5 & 100 & 0.0051 & 0.0998 & 0.0935 & 0.9300 \\ 
  GRTS-Model & Gridded & Normal & 0.5 & 100 & 0.0039 & 0.0982 & 0.0990 & 0.9470 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 17.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.5 & 100 & -0.0021 & 0.1211 & 0.1137 & 0.9445 \\ 
  IRS-Model & Gridded & Lognormal & 0.5 & 100 & -0.0058 & 0.1153 & 0.1032 & 0.9295 \\ 
  GRTS-Design & Gridded & Lognormal & 0.5 & 100 & 0.0041 & 0.1090 & 0.0910 & 0.9085 \\ 
  GRTS-Model & Gridded & Lognormal & 0.5 & 100 & 0.0019 & 0.1090 & 0.0959 & 0.9180 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 18.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.5 & 100 & -0.0064 & 0.1222 & 0.1190 & 0.9440 \\ 
  IRS-Model & Random & Normal & 0.5 & 100 & -0.0049 & 0.1073 & 0.1051 & 0.9440 \\ 
  GRTS-Design & Random & Normal & 0.5 & 100 & 0.0013 & 0.1041 & 0.0932 & 0.9155 \\ 
  GRTS-Model & Random & Normal & 0.5 & 100 & 0.0014 & 0.1020 & 0.0986 & 0.9400 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 19.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.5 & 100 & -0.0063 & 0.1204 & 0.1108 & 0.9355 \\ 
  IRS-Model & Random & Lognormal & 0.5 & 100 & -0.0097 & 0.1150 & 0.1005 & 0.9275 \\ 
  GRTS-Design & Random & Lognormal & 0.5 & 100 & 0.0003 & 0.1092 & 0.0891 & 0.8960 \\ 
  GRTS-Model & Random & Lognormal & 0.5 & 100 & -0.0004 & 0.1088 & 0.0937 & 0.9120 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 20.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.9 & 100 & -0.0007 & 0.1059 & 0.1078 & 0.9605 \\ 
  IRS-Model & Gridded & Normal & 0.9 & 100 & -0.0019 & 0.0700 & 0.0686 & 0.9445 \\ 
  GRTS-Design & Gridded & Normal & 0.9 & 100 & 0.0044 & 0.0655 & 0.0637 & 0.9435 \\ 
  GRTS-Model & Gridded & Normal & 0.9 & 100 & 0.0030 & 0.0585 & 0.0594 & 0.9440 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 21.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.9 & 100 & -0.0011 & 0.1092 & 0.0981 & 0.9540 \\ 
  IRS-Model & Gridded & Lognormal & 0.9 & 100 & -0.0033 & 0.0779 & 0.0674 & 0.9330 \\ 
  GRTS-Design & Gridded & Lognormal & 0.9 & 100 & 0.0035 & 0.0757 & 0.0621 & 0.9190 \\ 
  GRTS-Model & Gridded & Lognormal & 0.9 & 100 & 0.0017 & 0.0693 & 0.0585 & 0.9190 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 22.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.9 & 100 & -0.0048 & 0.1095 & 0.1058 & 0.9430 \\ 
  IRS-Model & Random & Normal & 0.9 & 100 & -0.0017 & 0.0671 & 0.0674 & 0.9535 \\ 
  GRTS-Design & Random & Normal & 0.9 & 100 & -0.0009 & 0.0674 & 0.0627 & 0.9265 \\ 
  GRTS-Model & Random & Normal & 0.9 & 100 & 0.0000 & 0.0595 & 0.0580 & 0.9420 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 23.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.9 & 100 & -0.0045 & 0.1110 & 0.0944 & 0.9375 \\ 
  IRS-Model & Random & Lognormal & 0.9 & 100 & -0.0052 & 0.0771 & 0.0645 & 0.9310 \\ 
  GRTS-Design & Random & Lognormal & 0.9 & 100 & -0.0009 & 0.0732 & 0.0596 & 0.9195 \\ 
  GRTS-Model & Random & Lognormal & 0.9 & 100 & -0.0007 & 0.0661 & 0.0555 & 0.9165 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 24.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0 & 200 & 0.0013 & 0.0882 & 0.0882 & 0.9550 \\ 
  IRS-Model & Gridded & Normal & 0 & 200 & 0.0013 & 0.0886 & 0.0880 & 0.9530 \\ 
  GRTS-Design & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.0854 & 0.9380 \\ 
  GRTS-Model & Gridded & Normal & 0 & 200 & -0.0045 & 0.0885 & 0.0874 & 0.9460 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 25.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0 & 200 & 0.0001 & 0.0880 & 0.0869 & 0.9350 \\ 
  IRS-Model & Gridded & Lognormal & 0 & 200 & -0.0002 & 0.0883 & 0.0866 & 0.9355 \\ 
  GRTS-Design & Gridded & Lognormal & 0 & 200 & -0.0034 & 0.0898 & 0.0843 & 0.9285 \\ 
  GRTS-Model & Gridded & Lognormal & 0 & 200 & -0.0035 & 0.0900 & 0.0862 & 0.9335 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 26.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0 & 200 & -0.0014 & 0.0893 & 0.0882 & 0.9465 \\ 
  IRS-Model & Random & Normal & 0 & 200 & -0.0015 & 0.0896 & 0.0880 & 0.9465 \\ 
  GRTS-Design & Random & Normal & 0 & 200 & 0.0007 & 0.0868 & 0.0855 & 0.9460 \\ 
  GRTS-Model & Random & Normal & 0 & 200 & 0.0007 & 0.0867 & 0.0873 & 0.9490 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 27.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0 & 200 & -0.0028 & 0.0891 & 0.0872 & 0.9425 \\ 
  IRS-Model & Random & Lognormal & 0 & 200 & -0.0033 & 0.0896 & 0.0870 & 0.9395 \\ 
  GRTS-Design & Random & Lognormal & 0 & 200 & 0.0015 & 0.0860 & 0.0840 & 0.9365 \\ 
  GRTS-Model & Random & Lognormal & 0 & 200 & 0.0012 & 0.0861 & 0.0857 & 0.9415 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 28.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.5 & 200 & 0.0022 & 0.0788 & 0.0796 & 0.9535 \\ 
  IRS-Model & Gridded & Normal & 0.5 & 200 & 0.0008 & 0.0678 & 0.0687 & 0.9580 \\ 
  GRTS-Design & Gridded & Normal & 0.5 & 200 & -0.0024 & 0.0671 & 0.0646 & 0.9335 \\ 
  GRTS-Model & Gridded & Normal & 0.5 & 200 & -0.0030 & 0.0661 & 0.0651 & 0.9410 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 29.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.5 & 200 & 0.0016 & 0.0816 & 0.0759 & 0.9420 \\ 
  IRS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0000 & 0.0739 & 0.0680 & 0.9335 \\ 
  GRTS-Design & Gridded & Lognormal & 0.5 & 200 & -0.0033 & 0.0744 & 0.0646 & 0.9330 \\ 
  GRTS-Model & Gridded & Lognormal & 0.5 & 200 & -0.0041 & 0.0742 & 0.0650 & 0.9350 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 30.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.5 & 200 & -0.0016 & 0.0790 & 0.0790 & 0.9480 \\ 
  IRS-Model & Random & Normal & 0.5 & 200 & -0.0010 & 0.0690 & 0.0684 & 0.9475 \\ 
  GRTS-Design & Random & Normal & 0.5 & 200 & -0.0002 & 0.0652 & 0.0644 & 0.9455 \\ 
  GRTS-Model & Random & Normal & 0.5 & 200 & 0.0001 & 0.0640 & 0.0647 & 0.9500 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 31.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.5 & 200 & -0.0027 & 0.0809 & 0.0738 & 0.9494 \\ 
  IRS-Model & Random & Lognormal & 0.5 & 200 & -0.0037 & 0.0732 & 0.0662 & 0.9454 \\ 
  GRTS-Design & Random & Lognormal & 0.5 & 200 & -0.0008 & 0.0671 & 0.0622 & 0.9434 \\ 
  GRTS-Model & Random & Lognormal & 0.5 & 200 & -0.0012 & 0.0662 & 0.0625 & 0.9479 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 32.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Normal & 0.9 & 200 & 0.0021 & 0.0704 & 0.0716 & 0.9570 \\ 
  IRS-Model & Gridded & Normal & 0.9 & 200 & 0.0005 & 0.0411 & 0.0417 & 0.9590 \\ 
  GRTS-Design & Gridded & Normal & 0.9 & 200 & -0.0005 & 0.0423 & 0.0410 & 0.9370 \\ 
  GRTS-Model & Gridded & Normal & 0.9 & 200 & -0.0012 & 0.0377 & 0.0370 & 0.9450 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 33.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Gridded & Lognormal & 0.9 & 200 & 0.0014 & 0.0744 & 0.0655 & 0.9520 \\ 
  IRS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0002 & 0.0480 & 0.0418 & 0.9355 \\ 
  GRTS-Design & Gridded & Lognormal & 0.9 & 200 & -0.0014 & 0.0499 & 0.0409 & 0.9380 \\ 
  GRTS-Model & Gridded & Lognormal & 0.9 & 200 & -0.0022 & 0.0459 & 0.0373 & 0.9380 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 34.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Normal & 0.9 & 200 & -0.0009 & 0.0725 & 0.0702 & 0.9470 \\ 
  IRS-Model & Random & Normal & 0.9 & 200 & -0.0010 & 0.0409 & 0.0409 & 0.9450 \\ 
  GRTS-Design & Random & Normal & 0.9 & 200 & -0.0005 & 0.0405 & 0.0401 & 0.9490 \\ 
  GRTS-Model & Random & Normal & 0.9 & 200 & 0.0001 & 0.0354 & 0.0358 & 0.9510 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 35.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrrr}
  \hline
Approach & Layout & Response & PDE & n & Bias & rMS(P)E & MStdE & Coverage \\ 
  \hline
IRS-Design & Random & Lognormal & 0.9 & 200 & -0.0006 & 0.0750 & 0.0626 & 0.9393 \\ 
  IRS-Model & Random & Lognormal & 0.9 & 200 & -0.0025 & 0.0461 & 0.0397 & 0.9418 \\ 
  GRTS-Design & Random & Lognormal & 0.9 & 200 & -0.0013 & 0.0452 & 0.0387 & 0.9348 \\ 
  GRTS-Model & Random & Lognormal & 0.9 & 200 & -0.0012 & 0.0397 & 0.0347 & 0.9418 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (PDE), sample size (n), mean (prediction) bias (Bias), root-mean-squared-(prediction) error (rMS(P)E), mean standard error (MStdE), and 95\% interval coverage (Coverage) in simulation scenario 36.} 
\end{table}

\end{document}