### 2022-04-11 inst/manuscript/manuscript.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Methods in Ecology and Evolution} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption} \usepackage{setspace} \usepackage{fancyhdr}

\pagestyle{fancy} \fancyhf{} \chead{Spatial design-based vs model-based}
\doublespacing

\begin{document}

\begin{frontmatter}

  \title{A comparison of design-based and model-based approaches for finite
population spatial data.}
    \author[USEPA]{Michael Dumelle\corref{1}}

    \author[STLAW]{Matt Higham}

    \author[NOAA]{Jay M. Ver Hoef}

    \author[USEPA]{Anthony R. Olsen}

    \author[OSU]{Lisa Madsen}

      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
Corvallis, Oregon, 97333}
    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
and Statistics, 23 Romoda Drive, Canton, New York, 13617}
    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
Corvallis, Oregon, 97331}
      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}

  \begin{abstract}
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    The design-based and model-based approaches to frequentist statistical
    inference rest on fundamentally different foundations. In the
    design-based approach, inference relies on random sampling. In the
    model-based approach, inference relies on distributional assumptions.
    We compare the approaches for finite population spatial data.
  \item
    We provide relevant background for the design-based and model-based
    approaches and then study their performance using simulated and real
    data. In the simulated data, a variety of sample sizes, location
    layouts, dependence structures, and response types are considered. In
    the simulated and real data, the population mean is the parameter of
    interest and performance is measured using statistics like bias,
    squared error, and interval coverage.
  \item
    When studying the simulated and real data, we found that regardless of
    the strength of spatial dependence in the data, the Generalized Random
    Tessellation Stratified (GRTS) algorithm, which explicitly
    incorporates spatial locations into sampling, tends to outperform the
    Simple Random Sampling (SRS) algorithm, which does not explicitly
    incorporate spatial locations into sampling. We also found that
    model-based approaches tend to outperform design-based approaches,
    even for skewed data where the model-based distributional assumptions
    are violated. The performance gap between these approaches is small
    GRTS samples are used but large when SRS samples are used. This
    suggests that the sampling choice (whether to use GRTS or SRS) is most
    important when performing design-based inference.
  \item
    There are many benefits and drawbacks to the design-based and
    model-based approaches for finite population spatial data that
    practitioners must consider when choosing between them. We provide
    relevant background contextualizing each approach and study their
    properties in a variety of scenarios, making recommendations for use
    based on the practitioner's goals.
  \end{enumerate}
  \end{abstract}

 \end{frontmatter}

\hypertarget{keywords}{%
\section*{Keywords}\label{keywords}}
\addcontentsline{toc}{section}{Keywords}

Design-based inference; Finite Population Block Kriging (FPBK);
Generalized Random Tessellation Stratified (GRTS) algorithm; Local
neighborhood variance estimator; Model-based inference; Restricted
Maximum Likelihood (REML) estimation; Spatially balanced sampling;
Spatial covariance

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

When data cannot be collected for all units in a population (i.e.,
population units), data are collected on a subset of the population
units -- this subset is called a sample. There are two general
approaches for using samples to make frequentist statistical inferences
about a population: design-based and model-based. In the design-based
approach, inference relies on randomly assigning some population units
to be in the sample (e.g., random sampling). Alternatively, in the
model-based approach, inference relies on distributional assumptions
about the underlying stochastic process that generated the sample. Each
paradigm has a deep historical context (Sterba, 2009) and its own set of
benefits and drawbacks (Hansen et al., 1983). In this manuscript, we
compare the design-based and model-based approaches for finite
population spatial data.

Spatial data are data that incorporate the locations of the population
units into either the sampling or estimation process. De Gruijter and
Ter Braak (1990) and Brus and DeGruijter (1993) give early comparisons
of design-based and model-based approaches for spatial data, quashing
the belief that design-based approaches could not be used for spatially
correlated data. Since then, there have been several general comparisons
between design-based and model-based approaches for spatial data (Brus
and De Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008; Wang et al.,
2012). Cooper (2006) reviews the two approaches in an ecological context
before introducing a ``model-assisted'' variance estimator that combines
aspects from each approach. In addition to Cooper (2006), there has been
substantial research and development into estimators that use both
design-based and model-based principles (see e.g., Sterba (2009) and
Cicchitelli and Montanari (2012), and see Chan-Golston et al. (2020) for
a Bayesian approach).

Certainly comparisons between design-based and model-based approaches
have been studied in spatial contexts. Our contribution is comparing
design-based approaches that incorporate spatial locations into sampling
and analysis to model-based approaches. Though the broad comparisons we
draw between design-based and model-based approaches generalize to
finite and infinite populations, we focus on finite populations. A
finite population contains a finite number of population units (we
assume the finite number is known); an example is lakes (treated as a
whole with the lake centroid representing location) in the contiguous
United States. An infinite population contains an infinite number of
population units; an example is locations within a single lake.

The rest of the manuscript is organized as follows. In Section
\ref{sec:background}, we introduce and provide relevant background for
the design-based and model-based approaches to finite population spatial
data. In Section \ref{sec:mm}, we describe how we compare performance of
the approaches with a simulation study and an analysis of real data that
contains mercury concentration in lakes located in the contiguous United
States. In Section \ref{sec:results}, we present results from the
simulation study and the mercury concentration analysis. And in Section
\ref{sec:discussion}, we end with a discussion and provide directions
for future research.

\hypertarget{sec:background}{%
\subsection{Background}\label{sec:background}}

The design-based and model-based approaches incorporate randomness in
fundamentally different ways. In this section, we describe the role of
randomness for each approach and the subsequent effects on statistical
inferences for spatial data.

\hypertarget{subsec:dvm_compare}{%
\subsubsection{Comparing Design-Based and Model-Based
Approaches}\label{subsec:dvm_compare}}

The design-based approach assumes the population is fixed. Randomness is
incorporated via the selection of population units according to a
sampling design. A sampling design assigns a non-zero probability of
inclusion (inclusion probability) in the sample to each population unit.
These inclusion probabilities are later used to estimate population
parameters. Some examples of commonly used sampling designs include
simple random sampling, stratified random sampling, and cluster
sampling.

When samples are chosen in a manner such that the layout of sampled
units reflects the layout of the population units, we call the resulting
sample ``spatially balanced.'' By ``reflecting the layout of the
population units'', we mean that if population units are concentrated in
specific areas, the units in the sample should be concentrated in the
same areas. Because spatially balanced samples reflect the layout of the
population units, they are not necessarily ``spread out'' in space in
some equidistant manner.

One approach to selecting spatially balanced samples is the Generalized
Random Tessellation Stratified (GRTS) algorithm (Stevens and Olsen,
2004), which we discuss in more detail in Section
\ref{subsec:spb_design}. When sampling designs do not incorporate
spatial locations into sampling, we call the resulting samples
``non-spatially balanced.''

Fundamentally, the design-based approach combines the randomness of the
sampling design with the data collected via the sample to justify the
estimation and uncertainty quantification of fixed, unknown parameters
of a population (e.g., a population mean). Treating the data as fixed
and incorporating randomness through the sampling design yields
estimators having very few other assumptions. Confidence intervals for
these types of estimators are typically derived using limiting arguments
that incorporate all possible samples. Sample means, for example, are
asymptotically normal (Gaussian) by the Central Limit Theorem (under
some assumptions). If we repeatedly select samples from the population,
then 95\% of all 95\% confidence intervals constructed from a procedure
with appropriate coverage will contain the true fixed population mean.
Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
design-based approach.

The model-based approach assumes the sample is a random realization of a
data-generating stochastic process. Randomness is formally incorporated
through distributional assumptions on this process. Strictly speaking,
randomness need not be incorporated through random sampling, though
Diggle et al. (2010) warn against preferential sampling. Preferential
sampling occurs when the process generating the data locations and the
process being modeled are not independent of one another. To guard
against preferential sampling, model-based approaches often still
implement some form of random sampling. When model-based approaches
implement random sampling, the inclusion probabilities are ignored when
analyzing the sample (in contrast to the design-based approach, which
relies on these inclusion probabilities to analyze the sample).

Instead of estimating fixed, unknown population parameters, as in the
design-based approach, often the goal of model-based inference is to
predict a realized variable, or value. For example, suppose the realized
mean of all population units is the value of interest. Instead of a
fixed, unknown mean, we are predicting the value of the mean, a random
variable. Prediction intervals are then derived using assumptions of the
data-generating stochastic process. If we repeatedly generate response
values from the same process and select samples, then 95\% of all 95\%
prediction intervals constructed from a procedure with appropriate
coverage will contain their respective realized means. Cressie (1993)
and Schabenberger and Gotway (2017) provide thorough reviews of
model-based approaches for spatial data. In Fig. \ref{fig:fig1}, we
provide a visual comparison of the design-based and model-based
approaches (Ver Hoef (2002) and Brus (2021) provide similar figures).

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/dvm_comp.jpeg}
  \caption{A visual comparison of the design-based and model-based approaches. In the top row, the design-based approach is highlighted. There is one fixed population with nine population units and three random samples of size four (points circled are those sampled). The response values at each site are fixed, but we obtain different estimates for the mean response in each random sample. In the bottom row, the model-based approach is highlighted. There are three realizations of the same data-generating stochastic process that are all sampled at the same four locations. The data-generating stochastic process has a single mean, but the mean of the nine population units is different in each of the three realizations.}
  \label{fig:fig1}
\end{figure}

\hypertarget{subsec:spb_design}{%
\subsubsection{Spatially Balanced Design and
Analysis}\label{subsec:spb_design}}

We previously mentioned that the design-based approach can be used to
select spatially balanced samples. Spatially balanced samples are useful
because parameter estimates from these samples tend to vary less than
parameter estimates from samples that are not spatially balanced
(Barabesi and Franceschi, 2011; Benedetti et al., 2017; Grafström and
Lundström, 2013; Robertson et al., 2013; Stevens and Olsen, 2004; Wang
et al., 2013). The first spatially balanced sampling algorithm to see
widespread use was the Generalized Random Tessellation Stratified (GRTS)
algorithm (Stevens and Olsen, 2004). To quantify the spatial balance of
a sample, Stevens and Olsen (2004) proposed loss metrics based on
Voronoi polygons (Dirichlet Tessellations). After the GRTS algorithm was
developed, several other spatially balanced sampling algorithms emerged,
including the Local Pivotal Method (Grafström et al., 2012; Grafström
and Matei, 2018), Spatially Correlated Poisson Sampling (Grafström,
2012), Balanced Acceptance Sampling (Robertson et al., 2013),
Within-Sample-Distance Sampling (Benedetti and Piersimoni, 2017), and
Halton Iterative Partitioning Sampling (Robertson et al., 2018). In this
manuscript, we select spatially balanced samples using the Generalized
Random Tessellation Stratified (GRTS) algorithm because it is readily
available in the \texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle
et al., 2022) and naturally accommodates finite and infinite sampling
frames, unequal inclusion probabilities, and replacement units
(replacement units are population units that can be sampled when a
population unit originally selected can no longer be sampled).

The GRTS algorithm selects samples by utilizing a particular mapping
between two-dimensional and one-dimensional space that preserves
proximity relationships. First the bounding box of the domain is split
up into four distinct, equally sized squares called level-one cells.
Each level-one is randomly assigned an level-one address of 0, 1, 2, or
3. The set of level-one cells is denoted by \(\mathcal{A}_1\) and
defined as \(\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}\). Within
each level-one cell, the inclusion probability for each population unit
is summed, and if any of these sums exceed one, a second level of cells
is added. Then each level-one cell is split into four distinct, equally
sized squares called level-two cells. Each level-two cell is randomly
assigned a level-two address of 0, 1, 2, or 3. The set of level-two
cells is denoted by \(\mathcal{A}_2\) and defined as
\(\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}\).
The inclusion probabilities within each level-two cell are summed, and
if any of these sums exceed one, a third level of cells is added. This
process continues for \(k\) steps, until all level-\(k\) cells have
inclusion probability sums no larger than one. Then
\(\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}\).

After determining \(\mathcal{A}_k\), it is placed into hierarchical
order. Hierarchical order is a numeric order that first sorts
\(\mathcal{A}_k\) by the level-one addresses from smallest to largest,
then sorts \(\mathcal{A}_k\) by the level-two addresses from smallest to
largest, and so on. For example, \(\mathcal{A}_2\) in hierarchical order
is the set
\(\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}\). Because
hierarchical ordering sorts by level-one cells, then level-two cells,
and so on, population units that have similar hierarchical addresses
tend to be nearby one another in space.Next each population unit is
mapped to a one-dimensional line in hierarchical order where each
population unit's inclusion probability equals its line-length. If a
level-\(k\) cell has multiple population units in it, they are randomly
placed within the cell's respective line segment. A uniform random
variable is then simulated in \([0, 1]\) and a systematic sample is
selected on the line, yielding \(n\) sample points for a sample size
\(n\). Each element in this systematic sample falls on some population
unit's line segment, and thus that population unit is selected in the
sample. For further details regarding the GRTS algorithm, see Stevens
and Olsen (2004).

After selecting a sample and collecting data, unbiased estimates of
population means and totals can be obtained using the Horvitz-Thompson
estimator (Horvitz and Thompson, 1952). If \(\tau\) is a population
total, the Horvitz-Thompson estimator for \(\tau\), denoted by
\(\hat{\tau}_{ht}\), is is given by \begin{align}\label{eq:ht}
  \hat{\tau}_{ht} = \sum_{i = 1}^n Z_i \pi_i^{-1},
\end{align} where \(Z_i\) is the value of the \(i\)th population unit in
the sample, \(\pi_i\) is the inclusion probability of the \(i\)th
population unit in the sample, and \(n\) is the sample size. An estimate
of the population mean is obtained by dividing \(\hat{\tau}_{ht}\) by
\(N\), the number of population units.

It is also important to quantify the uncertainty in \(\hat{\tau}_{ht}\).
Horvitz and Thompson (1952) and Sen (1953) provide variance estimators
for \(\hat{\tau}_{ht}\), but these estimators have two drawbacks. First,
they rely on calculating \(\pi_{ij}\), the probability that population
unit \(i\) and population unit \(j\) are both in the sample -- this
quantity can be challenging if not impossible to calculate analytically
for GRTS samples. Second, these estimators ignore the spatial locations
of the population units. To address these two drawbacks simultaneously,
Stevens and Olsen (2003) proposed the local neighborhood variance
estimator. The local neighborhood variance estimator does not rely on
\(\pi_{ij}\) and estimates the variance of \(\hat{\tau}\) conditional on
the random properties of the GRTS sample -- the idea being that this
conditioning should yield a more precise estimate of \(\hat{\tau}\).
They show that the each observation's contribution to the overall
variance is dominated by local variation. Thus the local neighborhood
variance estimator is a weighted sum of variance estimates from each
observation's local neighborhood. These local neighborhoods contain
observation itself and its three nearest neighbors. For more details,
see Stevens and Olsen (2003).

\hypertarget{finite-population-block-kriging}{%
\subsubsection{Finite Population Block
Kriging}\label{finite-population-block-kriging}}

Finite Population Block Kriging (FPBK) is a model-based approach that
expands the geostatistical Kriging framework to the finite population
setting (Ver Hoef, 2008). Instead of developing inference based on a
specific sampling design, we assume the data are generated by a spatial
stochastic process. We summarize some of the basic principles of FPBK
next -- for technical details, see Ver Hoef (2008). Let
\({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
be an \(N \times 1\) response vector at locations \(s_1\), \(s_2\), . .
. , \(s_N\) that can be measured at the \(N\) population units. Suppose
we want to use a sample to predict some linear function of the response
variable, \(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
\(\mathbf{b}^\prime\) is a \(1 \times N\) vector of weights (e.g, the
population mean is represented by a weights vector whose elements all
equal \(1 / N\)). Denoting quantities that are part of the sampled
population units with a subscript \emph{s} and quantities that are part
of the unsampled population units with a subscript \emph{u}, let

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
design matrices for the sampled and unsampled population units,
respectively, \(\bm{\beta}\) is the parameter vector of fixed effects,
and \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
sampled and unsampled population units, respectively.

FPBK assumes \(\bm{\delta}\) in Equation\(~\)\ref{equation:Zmarginal}
has mean-zero and a spatial dependence structure that can be modeled
using a covariance function. This covariance function is commonly
assumed to be non-negative, second-order stationary (depending only on
the distance between population units), isotropic (independent of
direction), and decay with distance between population units (Cressie,
1993). Henceforth, it is implied that we have made these same
assumptions regarding \(\bm{\delta}\), though Chiles and Delfiner
(1999), pp.~80-93 discuss covariance functions that are not second-order
stationary, not isotropic, or not either. A variety of flexible
covariance functions can be used to model \(\bm{\delta}\) (Cressie,
1993); one example is the exponential covariance function (Cressie
(1993) provides a thorough list of spatial covariance functions). The
\(i,j\)th element of the exponential covariance matrix,
\(\mathop{\mathrm{{cov}}}(\bm{\delta})\), is \mbox{}
\begin{align}\label{equation:expcov}
\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = 
\begin{cases} 
\sigma^2_{1}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
\sigma^2_{1} + \sigma^2_2 & h_{i,j} = 0
\end{cases}
,
\end{align} where \(\sigma^2_{1}\) is the variance parameter that
quantifies the spatially dependent variability, \(\sigma^2_{2}\) is the
variance parameter the quantifies that spatially independent
variability, \(\phi\) is the range parameter measuring the
distance-decay rate of the covariance, and \(h_{i,j}\) is the Euclidean
distance between population units \(i\) and \(j\). In geostatistical
literature, \(\sigma^2_{1}\) is often called the partial sill and
\(\sigma^2_{2}\) is often called the nugget.

The parameters in Equation\(~\)\ref{equation:Zmarginal} can be estimated
using a variety of techniques, but we focus on using restricted maximum
likelihood (Harville, 1977; Patterson and Thompson, 1971; Wolfinger et
al., 1994). REML is preferred over maximum likelihood (ML) because ML
estimates can be badly biased for small sample sizes, due to the fact
that ML makes no adjustment for the simultaneous estimation of
\(\bm{\beta}\) and \(\bm{\delta}\) (Patterson and Thompson, 1971). Minus
twice the REML log-likelihood of the sampled sites is given by
\begin{equation}\label{equation:reml}
  \ln|\bm{\Sigma}| + (\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}})^T \bm{\Sigma}_{ss}^{-1}(\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}}) + \ln|\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s| + (n - p) \ln(2 \pi) ,
\end{equation} where
\(\bm{\tilde{\beta}} = (\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s)^{-1} \bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{z}_s\)
and \(\bm{\Sigma}_{ss}\) is the covariance matrix of the sampled sites.
Minimizing Equation\(~\)\ref{equation:reml} yields
\(\bm{\hat{\delta}}_{reml}\), the REML estimates of \(\bm{\delta}\).
Then \(\bm{\beta}_{reml}\), the REML estimate of \(\bm{\beta}\), is
given by
\((\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s\),
where \(\bm{\hat{\Sigma}}_{ss}\) is \(\bm{\Sigma}_{ss}\) evaluated at
\(\bm{\hat{\delta}}_{reml}\).

With the model formulation in Equation\(~\)\ref{equation:Zmarginal}, the
Best Linear Unbiased Predictor (BLUP) for \(f(\mathbf{b}'\mathbf{z})\)
and its prediction variance can be computed. While details of the
derivation are in Ver Hoef (2008), we note here that the predictor and
its variance are both moment-based, meaning that they do not rely on any
distributional assumptions. Distributional assumptions are used,
however, when constructing prediction intervals.

Other approaches, such as k-nearest-neighbors (Fix and Hodges, 1989; Ver
Hoef and Temesgen, 2013) and random forest (Breiman, 2001), among
others, could also be used to obtain predictions for a mean or total
from finite population spatial data. Compared to the k-nearest-neighbors
and random forest approach, we prefer FPBK because it is model-based and
relies on theoretically-based variance estimators leveraging the model's
spatial covariance structure, whereas k-nearest-neighbors and random
forests use ad-hoc variance estimators (Ver Hoef and Temesgen, 2013).
Additionally, Ver Hoef and Temesgen (2013) compared FPBK,
k-nearest-neighbors, and random forest in a variety of spatial data
contexts, and FPBK tended to perform best.

\hypertarget{sec:mm}{%
\section{Materials and Methods}\label{sec:mm}}

\hypertarget{sec:mm_sim}{%
\subsection{Simulated Data}\label{sec:mm_sim}}

We used a simulation study to investigate performance of four
sampling-analysis combinations. The first sampling-analysis combination
was IRS-Design. In IRS-Design, samples were selected with the
Independent Random Sampling (IRS) algorithm. The IRS algorithm ignores
the spatial locations of the population units, thus the IRS samples were
not spatially balanced. In IRS-Design, samples were analyzed using the
design-based approach via the Horvitz-Thompson mean estimator and an IRS
variance estimator that ignored the spatial locations of the units in
the sample. The second sampling-analysis combination was IRS-Model,
where samples were selected with the IRS algorithm and analyzed using
the model-based approach while estimating the covariance parameters
(\(\bm{\delta}\)) and fixed effects (\(\bm{\beta}\) using restricted
maximum likelihood (REML). The third sampling-analysis combination was
GRTS-Design, where samples were selected with the GRTS algorithm and
analyzed using the design-based approach via the Horvitz-Thompson mean
estimator and the local neighborhood variance estimator (which does
incorporate the spatial locations of the units in the sample). The
fourth and final sampling-analysis combination was GRTS-Model, where
samples were selected with the GRTS algorithm and analyzed using the
model-based approach while estimating the covariance parameters
(\(\bm{\delta}\)) and fixed effects (\(\bm{\beta}\) using restricted
maximum likelihood (REML). These sampling-analysis combinations are also
provided in Table \ref{tab:designanalysis}. Lastly we note that for both
the IRS and GRTS samples, equal inclusion probabilities were assumed for
all population units. When IRS assumes equal inclusion probabilities for
all population units, the algorithm is equivalent to simple random
sampling (SRS).

\begin{table}[ht]
\centering
\begin{tabular}{r|ll}
  \hline
 & Design & Model \\ 
  \hline
IRS & IRS-Design & IRS-Model \\ 
  GRTS & GRTS-Design & GRTS-Model \\ 
   \hline
\end{tabular}
\caption{\label{tab:designanalysis} Sampling-analysis combinations in the simulation study. The rows give the two types of sampling designs and the columns give the two types of analyses.} 
\end{table}

CHANGE LOGNORMAL VERBAGE TO SKEWED

Performance for the four sampling-analysis combinations was evaluated in
36 different simulation scenarios. The 36 scenarios resulted from the
crossing of three sample sizes, two location layouts (of the population
units), two response types, and three proportions of dependent random
error. The three sample sizes (\(n\)) were \(n = 50, n = 100,\) and
\(n = 200\). Samples were always selected from a population size (\(N\))
of \(N = 900\). The two location layouts were random and gridded.
Locations in the random layout were randomly generated inside the unit
square (\([0, 1] \times [0, 1]\)). Locations in the gridded layout were
placed on a fixed, equally spaced grid inside the unit square. The two
response types were normal and lognormal. For the normal response type,
the response was simulated using mean-zero random errors with the
exponential covariance (Equation\(~\)\ref{equation:expcov}) for varying
proportions of dependent random error. The proportion of dependent
random error is represented by
\(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where \(\sigma^2_1\) and
\(\sigma^2_2\) are the dependent random error variance (partial sill)
and independent random error variance (nugget) from
Equation\(~\)\ref{equation:expcov}, respectively. The total variance,
\(\sigma^2_1 + \sigma^2_2\), was always 2. The range was always
\(\sqrt{2} / 3\), chosen so that the correlation in the dependent random
error decayed to nearly zero at \(\sqrt{2}\), the largest possible
distance between two population units in the domain. For the lognormal
response type, the response was first simulated using the same approach
as for the normal response type, except that the total variance was
0.6931 instead of 2. The response was then exponentiated, yielding a
lognormal random variable whose total variance was 2. The lognormal
responses were used to evaluate performance of the sampling-analysis
approaches for data that were skewed (i.e., not normal).

\begin{table}[ht]
\centering
\begin{tabular}{r|lll}
   \hline
Sample Size (n) & 50 & 100 & 200 \\ 
  Location Layout & Random & Gridded & - \\ 
  Proportion of Dependent Error & 0 & 0.5 & 0.9 \\ 
  Response Type & Normal & Lognormal & - \\ 
   \hline
\end{tabular}
\caption{\label{tab:parmtab} Simulation scenario options. All combinations of sample size, location layout, response type, and proportion of dependent random error composed the 36 simulation scenarios. In each simualtion scenario, the total variance was 2.} 
\end{table}

In each of the 36 simulation scenarios, there were 2000 independent
simulation trials. In each trial, IRS and GRTS samples were selected and
then design-based and model-based analyses were used to estimate
(design-based) or predict (model-based) the mean and construct 95\%
confidence (design-based) or 95\% prediction (model-based) intervals.
Then we recorded the bias, squared error, standard error, and interval
coverage for all sampling-analysis combinations. After all 2000 trials,
we summarized the long-run performance of the combinations by
calculating mean bias, rMS(P)E (root-mean-squared error for the
design-based approaches and root-mean-squared-prediction error for the
model-based approaches), MStdE (mean standard error), and the proportion
of times the true mean is contained in its 95\% confidence
(design-based) or 95\% prediction (model-based) interval. The 95\%
intervals were constructed using the normal distribution. Justification
for this comes from the asymptotic normality of means via the Central
Limit Theorem (under some assumptions). Quantifying mean bias and
rMS(P)E is important because they help us understand how far (under
different loss metrics) the estimates (design-based) or predictions
(model-based) tend to be from the true mean. Quantifying MStdE is
important because it helps us understand how precise intervals tend to
be. Quantifying interval coverage is important because it helps us
understand how often our 95\% intervals actually contain the true mean.

The IRS algorithm, IRS variance estimator, GRTS algorithm, and local
neighborhood variance estimator are available in the \texttt{spsurvey}
\textbf{\textsf{R}} package (Dumelle et al., 2022). FPBK is available in
the \texttt{sptotal} \textbf{\textsf{R}} package (Higham et al., 2021).

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
  \caption{}
  \label{fig:symm_pop_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
  \caption{}
  \label{fig:skew_pop_hist}
\end{subfigure}
\caption{Histograms of an example population realization for the normal (a) and skewed (b) scenarios for the simulated data.}
\label{fig:sim_pops}
\end{figure}

\hypertarget{sec:mm_app}{%
\subsection{National Lakes Assessment Data}\label{sec:mm_app}}

The United States Environmental Protection Agency (USEPA), states, and
tribes periodically conduct National Aquatic Research Surveys (NARS) to
assess the water quality of various bodies of water in the contiguous
United States. One component of NARS is the National Lakes Assessment
(NLA), which measures various aspects of lake health and water quality
(USEPA, 2012). We will analyze mercury concentration data collected at
986 lakes from the 2012 NLA. Although we can calculate the true mean
mercury concentration values for these 986 lakes, here we will explore
whether or not we can obtain an adequately precise estimate
(design-based) or prediction (model-based) for the realized mean mercury
concentration if we sample only 100 of the 986 lakes. For each of the
four familiar sampling-analysis combinations (IRS-Design, IRS-Model,
GRTS-Design, and GRTS-Model), we estimate (design-based) or predict
(model-based) the mean mercury concentration and construct 95\%
intervals from this sample of 100 lakes and compare to the true mean
mercury concentration from all 986 lakes.

\hypertarget{sec:results}{%
\section{Results}\label{sec:results}}

\hypertarget{sec:r_sim}{%
\subsection{Simulated Data}\label{sec:r_sim}}

The mean bias was nearly zero for all four sampling-analysis
combinations in all 36 scenarios, so we omit a more detailed summary of
those results here. Tables for mean bias in all 36 simulation scenarios
are provided in the supporting information.

Fig. \ref{fig:rmspe_eff} shows the relative rMS(P)E of the four sampling
analysis combinations using the random location layout with
``IRS-Design'' as the baseline. The relative rMS(P)E is defined as
\begin{equation*}
\frac{\text{rMS(P)E of sampling-analysis combination}}{\text{rMS(P)E of IRS-Design}},
\end{equation*} When there is no spatial covariance (Fig.
\ref{fig:rmspe_eff}, ``Prop DE: 0'' row), the four sampling-analysis
combinations have approximately equal rMS(P)E and using the GRTS
algorithm or a model-based analysis does not result in much, if any,
loss in efficiency compared to IRS-Design. When there is spatial
covariance (Fig. \ref{fig:rmspe_eff}, ``Prop DE: 0.5'' and ``Prop DE:
0.9'' rows), GRTS-Model tends to have the lowest rMS(P)E, followed by
GRTS-Design, IRS-Model, and finally IRS-Design, though the difference in
relative rMS(P)E among GRTS-Model, GRTS-Design, and IRS-Model is
relatively small. As the strength of spatial covariance increases, the
gap in rMS(P)E between IRS-Design and the other sampling-analysis
combinations widens. Finally we note that when there is spatial
covariance, IRS-Model has a much lower rMS(P)E than IRS-Design,
suggesting that the poor design properties of IRS are largely mitigated
by the model-based analysis. These rMS(P)E conclusions are similar to
those observed in the grid location layout, so we omit a grid location
layout figure here. Tables for rMS(P)E in all 36 simulation scenarios
are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
  \caption{Relative rMS(P)E in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
  \label{fig:rmspe_eff}
\end{figure}

Fig. \ref{fig:mse_eff} shows the relative MStdE of the four
sampling-analysis combinations using the random location layout with
``IRS-Design'' as the baseline. The relative MStdE is defined as
\begin{equation*}
\frac{\text{MStdE of sampling-analysis combination}}{\text{MStdE of IRS-Design}},
\end{equation*} Many general takeaways regarding MStdE are similar to
general takeaways regarding rMS(P)E: there seems to be no benefit to
using IRS, even when there is no spatial covariance; as the strength of
spatial covariance increases, the gap in MStdE between IRS-Design and
the other sampling-analysis combinations widens; and IRS-Model
outperforms IRS-Design by a noticeable margin. These fact that the
rMS(P)E and MStdE findings are similar is not particularly surprising
because the mean bias for all sampling-analysis combinations was nearly
zero, thus rMS(P)E is driven by the standard error of the estimators
(design-based) or predictors (model-based). We do note that between
GRTS-Design and GRTS-Model, GRTS-Design had lower MStdE when there was
no spatial covariance or a medium amount of spatial covariance (Fig.
\ref{fig:mse_eff}, ``Prop DE: 0'' and ``Prop DE: 0.5'' rows), and
GRTS-Model had lower MStdE when there was a high amount of spatial
covariance (Fig. \ref{fig:mse_eff}, ``Prop DE: 0.9'' row). These MStdE
conclusions are similar to those observed in the grid location layout,
so we omit a grid location layout figure here. Tables for MStdE in all
36 simulation scenarios are provided in the supporting information.

Fig. \ref{fig:figconf} shows the 95\% interval coverage for each of the
four sampling-analysis combinations in the random location layout.
Within each scenario, the sampling-analysis combinations tend to have
fairly similar interval coverage, though when \(n = 50\) or \(n = 100\),
GRTS-Design coverage is usually a few percentage points lower than the
other combinations. Coverage in the normal response scenarios was
usually near 95\%, while coverage in the lognormal response scenarios
usually varied from 90\% to 95\% but increased with the sample size. At
a sample size of 200, all four sampling-analysis combinations had
approximately 95\% interval coverage in both response scenarios for all
dependent error proportions. These interval coverage conclusions are
similar to those observed in the grid location layout, so we omit a grid
location layout figure here. Tables for interval coverage in all 36
simulation scenarios are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
  \caption{Interval coverage in the simulation study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black line represents 95\% coverage.}
  \label{fig:figconf}
\end{figure}

\hypertarget{sec:r_app}{%
\subsection{National Lakes Assessment DAta}\label{sec:r_app}}

Fig. \ref{fig:merc} shows a map and histogram of mercury concentration
in all 986 NLA lakes. The map shows mercury concentration exhibits some
spatial patterning, with high mercury concentrations in the northeast
and north central United States. The histogram shows that mercury
concentration is right-skewed, with most lakes having a low value of
mercury concentration but a few having a much higher concentration. Fig.
\ref{fig:merc} also shows mercury concentration's empirical
semivariogram. The empirical semivariogram can be used as a tool to
visualize spatial dependence. It quantifies the mean of the halved
squared differences (semivariance) among all pairs of mercury
concentrations at different distances apart. When a process has spatial
covariance (exhibits spatial dependence), the mean semivariance tends to
be smaller at small distances and larger at large distances. The
empirical semivariogram in Fig. \ref{fig:merc} suggests that mercury
concentration exhibits spatial dependence. Lastly we note that the true
mean mercury concentration in the 986 NLA lakes is 103.2 ng / g.

\begin{figure}
\centering
\begin{subfigure}{0.98\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
  \caption*{}
  \label{fig:mercury_map}
\end{subfigure} \\
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
  \caption*{}
  \label{fig:mercury_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/sv_plot.jpeg}
  \caption*{}
  \label{fig:sv_plot}
\end{subfigure}
\caption{Mercury concentration (Hg) visualizations for all 986 lakes in the NLA data. A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
\label{fig:merc}
\end{figure}

We selected a single IRS sample and a single GRTS sample and estimated
(design-based) or predicted (model-based) the mean mercury concentration
and constructed 95\% confidence (design-based) and 95\% (model-based)
prediction intervals. For the model-based analyses, the exponential
covariance was used. Table \ref{tab:appliedtab} shows the results from
these analyses. Though we should not generalize these results to other
samples from this population, we do mention a few findings. First,
IRS-Design has the largest standard error. Second, compared to
IRS-Design and IRS-Model, GRTS-Design and GRTS-Model are much closer to
the true mean mercury concentration (have bias closer to zero) and have
much lower standard errors (more precise intervals). Third, GRTS-Model
has the least amount of bias and the lowest standard error (most precise
interval). Finally, we note that for all sampling-analysis combinations,
the true mean mercury concentration (103.2 ng / g) is within the bounds
of the combination's 95\% interval.

\begin{table}[ht]
\centering
\begin{tabular}{lrrrrr}
  \hline
Approach & True Mean & Est/Pred & SE & 95\% LB & 95\% UB \\ 
  \hline
IRS-Design & 103.2 & 112.7 & 8.8 & 95.4 & 129.9 \\ 
  IRS-Model & 103.2 & 110.5 & 7.9 & 95.0 & 125.9 \\ 
  GRTS-Design & 103.2 & 101.8 & 6.1 & 89.8 & 113.7 \\ 
  GRTS-Model & 103.2 & 102.3 & 5.9 & 90.8 & 113.9 \\ 
   \hline
\end{tabular}
\caption{\label{tab:appliedtab} For each sampling-analysis combination (Approach), the true mean mercury concentration (True Mean), estimates/predictions (Est/Pred), standard errors (SE), lower 95\% interval bounds (95\% LB), and upper 95\% interval bounds (95\% UB) for mean mercury concentration computed using a sample of 100 lakes in the NLA data.} 
\end{table}

\hypertarget{new-application}{%
\subsection{New Application}\label{new-application}}

\begin{figure}
\centering
\begin{subfigure}{0.98\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
  \caption*{}
  \label{fig:zmmi_map}
\end{subfigure} \\
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
  \caption*{}
  \label{fig:zmmi_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/zmmi_sv_plot.jpeg}
  \caption*{}
  \label{fig:zmmi_sv_plot}
\end{subfigure}
\caption{zmmi visualizations for all 986 lakes in the NLA data. A spatial layout is in the top row, a histogram is in the bottom row and left column, and an empirical semivariogram is in the bottom row and right column.}
\label{fig:zmmi}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
  \caption{Relative rMS(P)E in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
  \label{fig:data_rmspe_eff}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/data_mse_eff.jpeg}
  \caption{Relative MStdE in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type.}
  \label{fig:data_mse_eff}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
  \caption{Interval coverage in the data study for the four sampling-analysis combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black line represents 95\% coverage.}
  \label{fig:data_figconf}
\end{figure}

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

The design-based and model-based approaches to statistical inference are
fundamentally different paradigms. The design-based approach relies on
random sampling to estimate population parameters. The model-based
approach relies on distributional assumptions to predict realized values
of a stochastic process. Though the model-based approach does not rely
on random sampling, it can still be beneficial as a way to guard against
preferential sampling. While the design-based and model-based approaches
have often been compared in the literature from theoretical and
analytical perspectives, our contribution lies in studying them in a
spatial context while implementing spatially balanced sampling and the
design-based, local neighborhood variance estimator. Aside from the
theoretical differences described, a few analytical findings from the
simulation study are particularly notable. First, independent of the
analysis approach, we found no reason to prefer IRS over GRTS when
sampling spatial data -- GRTS-Design and GRTS-Model generally had
similar rMS(P)E as their IRS counterparts when there was no spatial
covariance and lower rMS(P)E than their IRS counterparts when there was
spatial covariance. Second, the sampling decision (IRS vs GRTS) is most
important when using a design-based analysis. Though GRTS-Model still
had lower rMS(P)E than IRS-Model, the model-based analysis mitigated
most of the rMS(P)E inefficiencies that result from the IRS samples
lacking spatial balance. Third, as the strength of spatial covariance
increases, the gap in rMS(P)E and MStdE between IRS-Design and the other
sampling-analysis combinations also increases, likely because IRS-Design
is the only combination that ignores spatial locations in sampling and
analysis. Fourth and finally, when the response was normal, interval
coverage for all sampling-analysis combinations was usually close to
95\% for all sample sizes; when the response was lognormal, interval
coverage for all sampling-analysis combinations was usually between 90\%
and 95\% and closest to 95\% when \(n = 200\).

There are several benefits and drawbacks of the design-based and
model-based approaches for finite population spatial data. Some we have
discussed, but others we have not, and they are worthy of consideration
in future research. Design-based approaches are often computationally
efficient, while model-based approaches can be computationally
burdensome, especially for likelihood-based estimation methods like REML
that rely on inverting a covariance matrix. The design-based approach
easily handles binary data through a straightforward application of the
Horvitz-Thompson estimator. In contrast, analyzing binary data using a
model-based approach generally requires a logistic mixed regression
model, which can be challenging to estimate and interpret (Bolker et
al., 2009). The model-based approach, however, can more naturally
quantify the relationship between covariates (predictor variables) and
the response variable. The model-based approach also yields estimated
spatial covariance parameters, which help better understand the
dependence structure in the stochastic process of study. Model selection
is also possible using model-based approaches and criteria such as cross
validation, likelihood ratio tests, or AIC (Akaike, 1974). Model-based
approaches are capable of more efficient small-area estimation than
design-based approaches by leveraging distributional assumptions in
areas with few observed units. Model-based approaches can also compute
unit-by-unit predictions at unobserved locations and use them to
construct informative visualizations like smoothed maps. In short, when
deciding whether the design-based or model-based approach is more
appropriate to implement, the benefits and drawbacks of each approach
should be considered alongside the particular goals of the study.

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

We would like to thank the editors and anonymous reviewers for their
thoughtful comments which we feel greatly improved the manuscript.

The views expressed in this manuscript are those of the authors and do
not necessarily represent the views or policies of the U.S.
Environmental Protection Agency or the National Oceanic and Atmospheric
Administration. Any mention of trade names, products, or services does
not imply an endorsement by the U.S. government, the U.S. Environmental
Protection Agency, or the National Oceanic and Atmospheric
Administration. The U.S. Environmental Protection Agency and National
Oceanic and Atmospheric Administration do not endorse any commercial
products, services, or enterprises.

\hypertarget{conflict-of-interest-statement}{%
\section*{Conflict of Interest
Statement}\label{conflict-of-interest-statement}}
\addcontentsline{toc}{section}{Conflict of Interest Statement}

There are no conflicts of interest for any of the authors.

\hypertarget{author-contribution-statement}{%
\section*{Author Contribution
Statement}\label{author-contribution-statement}}
\addcontentsline{toc}{section}{Author Contribution Statement}

All authors conceived the ideas; All authors designed the methodology;
MD and MH performed the simulations and analyzed the data; MD and MH led
the writing of the manuscript; All authors contributed critically to the
drafts and gave final approval for publication.

\hypertarget{data-and-code-availability}{%
\section*{Data and Code Availability}\label{data-and-code-availability}}
\addcontentsline{toc}{section}{Data and Code Availability}

This manuscript has a supplementary \textbf{\textsf{R}} package that
contains all of the data and code used in its creation. The
supplementary \textbf{\textsf{R}} package is hosted on GitHub.
Instructions for download at available at

\url{https://github.com/michaeldumelle/DvMsp}.

If the manuscript is accepted, this repository will be archived in
Zenodo.

\hypertarget{supporting-information}{%
\section*{Supporting Information}\label{supporting-information}}
\addcontentsline{toc}{section}{Supporting Information}

In the supporting information, we provide tables of summary statistics
for all 36 simulation scenarios.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-akaike1974new}{}%
Akaike, H., 1974. A new look at the statistical model identification.
IEEE Transactions on Automatic Control 19, 716--723.

\leavevmode\hypertarget{ref-barabesi2011sampling}{}%
Barabesi, L., Franceschi, S., 2011. Sampling properties of spatial total
estimators under tessellation stratified designs. Environmetrics 22,
271--278.

\leavevmode\hypertarget{ref-benedetti2017spatially}{}%
Benedetti, R., Piersimoni, F., 2017. A spatially balanced design with
probability function proportional to the within sample distance.
Biometrical Journal 59, 1067--1084.

\leavevmode\hypertarget{ref-benedetti2017spatiallyreview}{}%
Benedetti, R., Piersimoni, F., Postiglione, P., 2017. Spatially balanced
sampling: A review and a reappraisal. International Statistical Review
85, 439--454.

\leavevmode\hypertarget{ref-bolker2009generalized}{}%
Bolker, B.M., Brooks, M.E., Clark, C.J., Geange, S.W., Poulsen, J.R.,
Stevens, M.H.H., White, J.-S.S., 2009. Generalized linear mixed models:
A practical guide for ecology and evolution. Trends in ecology \&
evolution 24, 127--135.

\leavevmode\hypertarget{ref-breiman2001random}{}%
Breiman, L., 2001. Random forests. Machine Learning 45, 5--32.

\leavevmode\hypertarget{ref-brus1997random}{}%
Brus, D., De Gruijter, J., 1997. Random sampling or geostatistical
modelling? Choosing between design-based and model-dased sampling
strategies for soil (with discussion). Geoderma 80, 1--44.

\leavevmode\hypertarget{ref-brus2021statistical}{}%
Brus, D.J., 2021. Statistical approaches for spatial sample survey:
Persistent misconceptions and new developments. European Journal of Soil
Science 72, 686--703.

\leavevmode\hypertarget{ref-brus1993design}{}%
Brus, D.J., DeGruijter, J.J., 1993. Design-based versus model-based
estimates of spatial means: Theory and application in environmental soil
science. Environmetrics 4, 123--152.

\leavevmode\hypertarget{ref-chan2020bayesian}{}%
Chan-Golston, A.M., Banerjee, S., Handcock, M.S., 2020. Bayesian
inference for finite populations under spatial process settings.
Environmetrics 31, e2606.

\leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
Chiles, J.-P., Delfiner, P., 1999. Geostatistics: Modeling Spatial
Uncertainty. John Wiley \& Sons, New York.

\leavevmode\hypertarget{ref-cicchitelli2012model}{}%
Cicchitelli, G., Montanari, G.E., 2012. Model-assisted estimation of a
spatial population mean. International Statistical Review 80, 111--126.

\leavevmode\hypertarget{ref-cooper2006sampling}{}%
Cooper, C., 2006. Sampling and variance estimation on continuous
domains. Environmetrics 17, 539--553.

\leavevmode\hypertarget{ref-cressie1993statistics}{}%
Cressie, N., 1993. Statistics for spatial data. John Wiley \& Sons.

\leavevmode\hypertarget{ref-de1990model}{}%
De Gruijter, J., Ter Braak, C., 1990. Model-free estimation from spatial
samples: A reappraisal of classical sampling theory. Mathematical
Geology 22, 407--415.

\leavevmode\hypertarget{ref-diggle2010geostatistical}{}%
Diggle, P.J., Menezes, R., Su, T.-l., 2010. Geostatistical inference
under preferential sampling. Journal of the Royal Statistical Society:
Series C (Applied Statistics) 59, 191--232.

\leavevmode\hypertarget{ref-dumelle2022spsurvey}{}%
Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2022. Spsurvey:
Spatial sampling design and analysis.

\leavevmode\hypertarget{ref-fix1989discriminatory}{}%
Fix, E., Hodges, J.L., 1989. Discriminatory analysis. Nonparametric
discrimination: Consistency properties. International Statistical
Review/Revue Internationale de Statistique 57, 238--247.

\leavevmode\hypertarget{ref-grafstrom2012spatiallypoisson}{}%
Grafström, A., 2012. Spatially correlated poisson sampling. Journal of
Statistical Planning and Inference 142, 139--147.

\leavevmode\hypertarget{ref-grafstrom2013well}{}%
Grafström, A., Lundström, N.L., 2013. Why well spread probability
samples are balanced. Open Journal of Statistics 3, 36--41.

\leavevmode\hypertarget{ref-grafstrom2012spatially}{}%
Grafström, A., Lundström, N.L., Schelin, L., 2012. Spatially balanced
sampling through the pivotal method. Biometrics 68, 514--520.

\leavevmode\hypertarget{ref-grafstrom2018spatially}{}%
Grafström, A., Matei, A., 2018. Spatially balanced sampling of
continuous populations. Scandinavian Journal of Statistics 45, 792--805.

\leavevmode\hypertarget{ref-hansen1983evaluation}{}%
Hansen, M.H., Madow, W.G., Tepping, B.J., 1983. An evaluation of
model-dependent and probability-sampling inferences in sample surveys.
Journal of the American Statistical Association 78, 776--793.

\leavevmode\hypertarget{ref-harville1977maximum}{}%
Harville, D.A., 1977. Maximum likelihood approaches to variance
component estimation and to related problems. Journal of the American
Statistical Association 72, 320--338.

\leavevmode\hypertarget{ref-higham2021sptotal}{}%
Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
Predicting totals and weighted sums from spatial data.

\leavevmode\hypertarget{ref-horvitz1952generalization}{}%
Horvitz, D.G., Thompson, D.J., 1952. A generalization of sampling
without replacement from a finite universe. Journal of the American
Statistical Association 47, 663--685.

\leavevmode\hypertarget{ref-lohr2009sampling}{}%
Lohr, S.L., 2009. Sampling: Design and analysis. Nelson Education.

\leavevmode\hypertarget{ref-patterson1971recovery}{}%
Patterson, H.D., Thompson, R., 1971. Recovery of inter-block information
when block sizes are unequal. Biometrika 58, 545--554.

\leavevmode\hypertarget{ref-robertson2013bas}{}%
Robertson, B., Brown, J., McDonald, T., Jaksons, P., 2013. BAS: Balanced
acceptance sampling of natural resources. Biometrics 69, 776--784.

\leavevmode\hypertarget{ref-robertson2018halton}{}%
Robertson, B., McDonald, T., Price, C., Brown, J., 2018. Halton
iterative partitioning: Spatially balanced sampling via partitioning.
Environmental and Ecological Statistics 25, 305--323.

\leavevmode\hypertarget{ref-sarndal2003model}{}%
Särndal, C.-E., Swensson, B., Wretman, J., 2003. Model assisted survey
sampling. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-schabenberger2017statistical}{}%
Schabenberger, O., Gotway, C.A., 2017. Statistical methods for spatial
data analysis. CRC press.

\leavevmode\hypertarget{ref-sen1953estimate}{}%
Sen, A.R., 1953. On the estimate of the variance in sampling with
varying probabilities. Journal of the Indian Society of Agricultural
Statistics 5, 127.

\leavevmode\hypertarget{ref-sterba2009alternative}{}%
Sterba, S.K., 2009. Alternative model-based and design-based frameworks
for inference from samples to populations: From polarization to
integration. Multivariate Behavioral Research 44, 711--740.

\leavevmode\hypertarget{ref-stevens2003variance}{}%
Stevens, D.L., Olsen, A.R., 2003. Variance estimation for spatially
balanced samples of environmental resources. Environmetrics 14,
593--610.

\leavevmode\hypertarget{ref-stevens2004spatially}{}%
Stevens, D.L., Olsen, A.R., 2004. Spatially balanced sampling of natural
resources. Journal of the American Statistical Association 99, 262--278.

\leavevmode\hypertarget{ref-USEPA2012NLA}{}%
USEPA, 2012. National lakes assessment 2012.
https://www.epa.gov/national-aquatic-resource-surveys/national-results-and-regional-highlights-national-lakes-assessment.

\leavevmode\hypertarget{ref-verhoef2002sampling}{}%
Ver Hoef, J., 2002. Sampling and geostatistics for spatial data.
Ecoscience 9, 152--161.

\leavevmode\hypertarget{ref-verhoef2008spatial}{}%
Ver Hoef, J.M., 2008. Spatial methods for plot-based sampling of
wildlife populations. Environmental and Ecological Statistics 15, 3--13.

\leavevmode\hypertarget{ref-ver2013comparison}{}%
Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
model to nearest neighbor (k-nn) methods for forestry applications. PlOS
ONE 8, e59129.

\leavevmode\hypertarget{ref-wang2013design}{}%
Wang, J.-F., Jiang, C.-S., Hu, M.-G., Cao, Z.-D., Guo, Y.-S., Li, L.-F.,
Liu, T.-J., Meng, B., 2013. Design-based spatial sampling: Theory and
implementation. Environmental Modelling \& Software 40, 280--288.

\leavevmode\hypertarget{ref-wang2012review}{}%
Wang, J.-F., Stein, A., Gao, B.-B., Ge, Y., 2012. A review of spatial
sampling. Spatial Statistics 2, 1--14.

\leavevmode\hypertarget{ref-wolfinger1994computing}{}%
Wolfinger, R., Tobias, R., Sall, J., 1994. Computing gaussian
likelihoods and their derivatives for general linear mixed models. SIAM
Journal on Scientific Computing 15, 1294--1310.

\end{document}