### 2022-04-13 inst/manuscript/supporting.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Methods in Ecology and Evolution} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial data -- Supporting information.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \renewenvironment{abstract}{}{}

\begin{document}

\begin{frontmatter}

  \title{A comparison of design-based and model-based approaches for finite
population spatial data -- Supporting information.}
    \author[USEPA]{Michael Dumelle\corref{1}}

    \author[STLAW]{Matt Higham}

    \author[OSU]{Lisa Madsen}

    \author[USEPA]{Anthony R. Olsen}

    \author[NOAA]{Jay M. Ver Hoef}

      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
Corvallis, Oregon, 97333}
    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
and Statistics, 23 Romoda Drive, Canton, New York, 13617}
    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
Corvallis, Oregon, 97331}
    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}

  \begin{abstract}

  \end{abstract}

 \end{frontmatter}

Recall the four sampling-analysis combinations from the simulated and
real data: simple random sampling with design-based inference (SRS-DB),
simple random sapling with model-based inference (SRS-MB), GRTS sampling
with design-based inference (GRTS-DB), and GRTS sampling with
model-based inference (GRTS-MB).

\hypertarget{sec:simtabs}{%
\section{Simulated Data}\label{sec:simtabs}}

For the simulated data, we considered 36 parameter configurations -- the
crossing of three sample sizes (\(n = 50\), \(n = 100\), \(n = 200)\),
two location layouts (random and gridded), three proportions of
dependent random error (DRE) (0\%, 50\%, 90\%), and two response types
(normal and skewed).

Next we present tables summarizing mean bias, RMSE, and interval
coverage for all 36 simulation scenarios.

\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 0\% & 50 & 0.0004 & 0.1922 & 0.9345 \\ 
  SRS-MB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1930 & 0.9420 \\ 
  GRTS-DB & Gridded & Normal & 0\% & 50 & 0.0005 & 0.1914 & 0.9170 \\ 
  GRTS-MB & Gridded & Normal & 0\% & 50 & -0.0112 & 0.1905 & 0.9490 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 1.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 0\% & 50 & 0.0004 & 0.1965 & 0.8955 \\ 
  SRS-MB & Gridded & Skewed & 0\% & 50 & -0.0110 & 0.2055 & 0.9045 \\ 
  GRTS-DB & Gridded & Skewed & 0\% & 50 & 0.0025 & 0.1947 & 0.8710 \\ 
  GRTS-MB & Gridded & Skewed & 0\% & 50 & -0.0048 & 0.1988 & 0.9110 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 2.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 0\% & 50 & -0.0002 & 0.1964 & 0.9295 \\ 
  SRS-MB & Random & Normal & 0\% & 50 & 0.0045 & 0.1953 & 0.9395 \\ 
  GRTS-DB & Random & Normal & 0\% & 50 & -0.0004 & 0.1951 & 0.9065 \\ 
  GRTS-MB & Random & Normal & 0\% & 50 & 0.0031 & 0.1940 & 0.9425 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 3.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 0\% & 50 & 0.0016 & 0.1941 & 0.8985 \\ 
  SRS-MB & Random & Skewed & 0\% & 50 & -0.0059 & 0.1932 & 0.9130 \\ 
  GRTS-DB & Random & Skewed & 0\% & 50 & 0.0041 & 0.1910 & 0.8740 \\ 
  GRTS-MB & Random & Skewed & 0\% & 50 & 0.0000 & 0.1861 & 0.9225 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 4.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 50\% & 50 & -0.0031 & 0.1472 & 0.9400 \\ 
  SRS-MB & Gridded & Normal & 50\% & 50 & -0.0007 & 0.1617 & 0.9355 \\ 
  GRTS-DB & Gridded & Normal & 50\% & 50 & -0.0030 & 0.1501 & 0.9205 \\ 
  GRTS-MB & Gridded & Normal & 50\% & 50 & -0.0013 & 0.1775 & 0.9460 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 5.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 50\% & 50 & -0.0060 & 0.1605 & 0.9085 \\ 
  SRS-MB & Gridded & Skewed & 50\% & 50 & -0.0083 & 0.1787 & 0.8990 \\ 
  GRTS-DB & Gridded & Skewed & 50\% & 50 & -0.0041 & 0.1610 & 0.8995 \\ 
  GRTS-MB & Gridded & Skewed & 50\% & 50 & -0.0047 & 0.1891 & 0.9100 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 6.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 50\% & 50 & -0.0097 & 0.1507 & 0.9315 \\ 
  SRS-MB & Random & Normal & 50\% & 50 & 0.0042 & 0.1615 & 0.9350 \\ 
  GRTS-DB & Random & Normal & 50\% & 50 & -0.0092 & 0.1513 & 0.9165 \\ 
  GRTS-MB & Random & Normal & 50\% & 50 & 0.0061 & 0.1744 & 0.9445 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 7.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 50\% & 50 & -0.0021 & 0.1624 & 0.9000 \\ 
  SRS-MB & Random & Skewed & 50\% & 50 & -0.0051 & 0.1808 & 0.9065 \\ 
  GRTS-DB & Random & Skewed & 50\% & 50 & 0.0021 & 0.1618 & 0.8880 \\ 
  GRTS-MB & Random & Skewed & 50\% & 50 & 0.0014 & 0.1804 & 0.9195 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 8.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 90\% & 50 & -0.0037 & 0.0960 & 0.9400 \\ 
  SRS-MB & Gridded & Normal & 90\% & 50 & 0.0044 & 0.1137 & 0.9435 \\ 
  GRTS-DB & Gridded & Normal & 90\% & 50 & -0.0021 & 0.1072 & 0.9280 \\ 
  GRTS-MB & Gridded & Normal & 90\% & 50 & 0.0055 & 0.1621 & 0.9355 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 9.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 90\% & 50 & -0.0039 & 0.1039 & 0.9205 \\ 
  SRS-MB & Gridded & Skewed & 90\% & 50 & -0.0011 & 0.1217 & 0.9130 \\ 
  GRTS-DB & Gridded & Skewed & 90\% & 50 & -0.0017 & 0.1129 & 0.9180 \\ 
  GRTS-MB & Gridded & Skewed & 90\% & 50 & 0.0007 & 0.1546 & 0.9345 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 10.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 90\% & 50 & -0.0011 & 0.0940 & 0.9410 \\ 
  SRS-MB & Random & Normal & 90\% & 50 & 0.0045 & 0.1093 & 0.9445 \\ 
  GRTS-DB & Random & Normal & 90\% & 50 & -0.0017 & 0.1054 & 0.9295 \\ 
  GRTS-MB & Random & Normal & 90\% & 50 & 0.0058 & 0.1575 & 0.9470 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 11.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1127 & 0.9225 \\ 
  SRS-MB & Random & Skewed & 90\% & 50 & -0.0046 & 0.1301 & 0.9180 \\ 
  GRTS-DB & Random & Skewed & 90\% & 50 & -0.0026 & 0.1250 & 0.9175 \\ 
  GRTS-MB & Random & Skewed & 90\% & 50 & -0.0022 & 0.1772 & 0.9305 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 12.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 0\% & 100 & -0.0010 & 0.1352 & 0.9445 \\ 
  SRS-MB & Gridded & Normal & 0\% & 100 & -0.0026 & 0.1362 & 0.9405 \\ 
  GRTS-DB & Gridded & Normal & 0\% & 100 & -0.0011 & 0.1345 & 0.9210 \\ 
  GRTS-MB & Gridded & Normal & 0\% & 100 & -0.0028 & 0.1352 & 0.9475 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 13.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 0\% & 100 & 0.0077 & 0.1307 & 0.9155 \\ 
  SRS-MB & Gridded & Skewed & 0\% & 100 & -0.0024 & 0.1412 & 0.9150 \\ 
  GRTS-DB & Gridded & Skewed & 0\% & 100 & 0.0082 & 0.1304 & 0.8945 \\ 
  GRTS-MB & Gridded & Skewed & 0\% & 100 & -0.0001 & 0.1384 & 0.9210 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 14.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 0\% & 100 & 0.0019 & 0.1318 & 0.9460 \\ 
  SRS-MB & Random & Normal & 0\% & 100 & -0.0007 & 0.1344 & 0.9435 \\ 
  GRTS-DB & Random & Normal & 0\% & 100 & 0.0021 & 0.1315 & 0.9280 \\ 
  GRTS-MB & Random & Normal & 0\% & 100 & -0.0004 & 0.1339 & 0.9450 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 15.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 0\% & 100 & 0.0039 & 0.1315 & 0.9230 \\ 
  SRS-MB & Random & Skewed & 0\% & 100 & 0.0005 & 0.1342 & 0.9280 \\ 
  GRTS-DB & Random & Skewed & 0\% & 100 & 0.0043 & 0.1311 & 0.8990 \\ 
  GRTS-MB & Random & Skewed & 0\% & 100 & 0.0021 & 0.1322 & 0.9280 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 16.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 50\% & 100 & -0.0003 & 0.0993 & 0.9435 \\ 
  SRS-MB & Gridded & Normal & 50\% & 100 & -0.0007 & 0.1058 & 0.9425 \\ 
  GRTS-DB & Gridded & Normal & 50\% & 100 & 0.0000 & 0.1003 & 0.9300 \\ 
  GRTS-MB & Gridded & Normal & 50\% & 100 & -0.0025 & 0.1207 & 0.9420 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 17.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 50\% & 100 & -0.0019 & 0.1127 & 0.9175 \\ 
  SRS-MB & Gridded & Skewed & 50\% & 100 & 0.0002 & 0.1135 & 0.9160 \\ 
  GRTS-DB & Gridded & Skewed & 50\% & 100 & 0.0000 & 0.1113 & 0.9025 \\ 
  GRTS-MB & Gridded & Skewed & 50\% & 100 & 0.0023 & 0.1221 & 0.9315 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 18.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 50\% & 100 & -0.0010 & 0.0997 & 0.9380 \\ 
  SRS-MB & Random & Normal & 50\% & 100 & -0.0001 & 0.1079 & 0.9440 \\ 
  GRTS-DB & Random & Normal & 50\% & 100 & -0.0012 & 0.1012 & 0.9235 \\ 
  GRTS-MB & Random & Normal & 50\% & 100 & -0.0016 & 0.1214 & 0.9440 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 19.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 50\% & 100 & -0.0002 & 0.1105 & 0.9235 \\ 
  SRS-MB & Random & Skewed & 50\% & 100 & -0.0020 & 0.1097 & 0.9290 \\ 
  GRTS-DB & Random & Skewed & 50\% & 100 & 0.0013 & 0.1104 & 0.9090 \\ 
  GRTS-MB & Random & Skewed & 50\% & 100 & 0.0020 & 0.1188 & 0.9395 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 20.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0602 & 0.9425 \\ 
  SRS-MB & Gridded & Normal & 90\% & 100 & -0.0003 & 0.0701 & 0.9440 \\ 
  GRTS-DB & Gridded & Normal & 90\% & 100 & -0.0013 & 0.0670 & 0.9325 \\ 
  GRTS-MB & Gridded & Normal & 90\% & 100 & -0.0027 & 0.1083 & 0.9460 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 21.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 90\% & 100 & 0.0004 & 0.0695 & 0.9280 \\ 
  SRS-MB & Gridded & Skewed & 90\% & 100 & -0.0038 & 0.0767 & 0.9265 \\ 
  GRTS-DB & Gridded & Skewed & 90\% & 100 & 0.0018 & 0.0746 & 0.9270 \\ 
  GRTS-MB & Gridded & Skewed & 90\% & 100 & -0.0015 & 0.1108 & 0.9345 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 22.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 90\% & 100 & -0.0007 & 0.0570 & 0.9480 \\ 
  SRS-MB & Random & Normal & 90\% & 100 & 0.0011 & 0.0675 & 0.9475 \\ 
  GRTS-DB & Random & Normal & 90\% & 100 & 0.0009 & 0.0651 & 0.9430 \\ 
  GRTS-MB & Random & Normal & 90\% & 100 & 0.0002 & 0.1071 & 0.9490 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 23.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 90\% & 100 & 0.0023 & 0.0671 & 0.9230 \\ 
  SRS-MB & Random & Skewed & 90\% & 100 & -0.0031 & 0.0757 & 0.9315 \\ 
  GRTS-DB & Random & Skewed & 90\% & 100 & 0.0026 & 0.0751 & 0.9225 \\ 
  GRTS-MB & Random & Skewed & 90\% & 100 & -0.0019 & 0.1165 & 0.9325 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 24.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0843 & 0.9610 \\ 
  SRS-MB & Gridded & Normal & 0\% & 200 & 0.0003 & 0.0894 & 0.9440 \\ 
  GRTS-DB & Gridded & Normal & 0\% & 200 & 0.0032 & 0.0842 & 0.9520 \\ 
  GRTS-MB & Gridded & Normal & 0\% & 200 & 0.0005 & 0.0891 & 0.9465 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 25.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 0\% & 200 & 0.0005 & 0.0892 & 0.9405 \\ 
  SRS-MB & Gridded & Skewed & 0\% & 200 & -0.0023 & 0.0895 & 0.9445 \\ 
  GRTS-DB & Gridded & Skewed & 0\% & 200 & 0.0006 & 0.0891 & 0.9360 \\ 
  GRTS-MB & Gridded & Skewed & 0\% & 200 & -0.0018 & 0.0894 & 0.9440 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 26.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 0\% & 200 & 0.0009 & 0.0882 & 0.9480 \\ 
  SRS-MB & Random & Normal & 0\% & 200 & 0.0017 & 0.0907 & 0.9375 \\ 
  GRTS-DB & Random & Normal & 0\% & 200 & 0.0008 & 0.0881 & 0.9400 \\ 
  GRTS-MB & Random & Normal & 0\% & 200 & 0.0019 & 0.0905 & 0.9385 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 27.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9395 \\ 
  SRS-MB & Random & Skewed & 0\% & 200 & 0.0005 & 0.0869 & 0.9390 \\ 
  GRTS-DB & Random & Skewed & 0\% & 200 & -0.0011 & 0.0873 & 0.9345 \\ 
  GRTS-MB & Random & Skewed & 0\% & 200 & 0.0007 & 0.0867 & 0.9400 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 28.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 50\% & 200 & 0.0039 & 0.0679 & 0.9375 \\ 
  SRS-MB & Gridded & Normal & 50\% & 200 & -0.0004 & 0.0678 & 0.9510 \\ 
  GRTS-DB & Gridded & Normal & 50\% & 200 & 0.0036 & 0.0688 & 0.9330 \\ 
  GRTS-MB & Gridded & Normal & 50\% & 200 & -0.0016 & 0.0777 & 0.9520 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 29.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 50\% & 200 & 0.0018 & 0.0691 & 0.9360 \\ 
  SRS-MB & Gridded & Skewed & 50\% & 200 & 0.0016 & 0.0747 & 0.9400 \\ 
  GRTS-DB & Gridded & Skewed & 50\% & 200 & 0.0019 & 0.0698 & 0.9325 \\ 
  GRTS-MB & Gridded & Skewed & 50\% & 200 & 0.0013 & 0.0816 & 0.9400 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 30.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 50\% & 200 & 0.0009 & 0.0643 & 0.9510 \\ 
  SRS-MB & Random & Normal & 50\% & 200 & 0.0005 & 0.0713 & 0.9444 \\ 
  GRTS-DB & Random & Normal & 50\% & 200 & 0.0005 & 0.0655 & 0.9474 \\ 
  GRTS-MB & Random & Normal & 50\% & 200 & 0.0008 & 0.0816 & 0.9389 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 31.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 50\% & 200 & -0.0007 & 0.0679 & 0.9385 \\ 
  SRS-MB & Random & Skewed & 50\% & 200 & -0.0048 & 0.0729 & 0.9350 \\ 
  GRTS-DB & Random & Skewed & 50\% & 200 & -0.0001 & 0.0680 & 0.9400 \\ 
  GRTS-MB & Random & Skewed & 50\% & 200 & -0.0041 & 0.0789 & 0.9510 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 32.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0376 & 0.9475 \\ 
  SRS-MB & Gridded & Normal & 90\% & 200 & -0.0008 & 0.0418 & 0.9445 \\ 
  GRTS-DB & Gridded & Normal & 90\% & 200 & -0.0010 & 0.0418 & 0.9455 \\ 
  GRTS-MB & Gridded & Normal & 90\% & 200 & -0.0011 & 0.0730 & 0.9450 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 33.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Gridded & Skewed & 90\% & 200 & 0.0010 & 0.0423 & 0.9375 \\ 
  SRS-MB & Gridded & Skewed & 90\% & 200 & -0.0007 & 0.0475 & 0.9365 \\ 
  GRTS-DB & Gridded & Skewed & 90\% & 200 & 0.0018 & 0.0482 & 0.9370 \\ 
  GRTS-MB & Gridded & Skewed & 90\% & 200 & 0.0038 & 0.0768 & 0.9430 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 34.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Normal & 90\% & 200 & 0.0014 & 0.0368 & 0.9475 \\ 
  SRS-MB & Random & Normal & 90\% & 200 & 0.0011 & 0.0408 & 0.9480 \\ 
  GRTS-DB & Random & Normal & 90\% & 200 & 0.0016 & 0.0418 & 0.9450 \\ 
  GRTS-MB & Random & Normal & 90\% & 200 & 0.0026 & 0.0713 & 0.9495 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 35.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{lllrrrrr}
  \hline
Approach & Layout & Response & DRE\% & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Random & Skewed & 90\% & 200 & -0.0001 & 0.0396 & 0.9419 \\ 
  SRS-MB & Random & Skewed & 90\% & 200 & 0.0011 & 0.0437 & 0.9414 \\ 
  GRTS-DB & Random & Skewed & 90\% & 200 & -0.0008 & 0.0431 & 0.9504 \\ 
  GRTS-MB & Random & Skewed & 90\% & 200 & 0.0028 & 0.0708 & 0.9504 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in simulation scenario 36.} 
\end{table}

\hypertarget{sec:datatabs}{%
\section{National Lakes Assessment (Real) Data}\label{sec:datatabs}}

Next we present tables summarizing mean bias, RMSE, and interval
coverage for ZMMI and Hg ppb data.

\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{llrrrr}
  \hline
Approach & Response & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Hg ppb & 50 & -0.5604 & 10.4434 & 0.9070 \\ 
  SRS-MB & Hg ppb & 50 & 0.0573 & 11.3760 & 0.9020 \\ 
  GRTS-DB & Hg ppb & 50 & -0.1880 & 10.5558 & 0.8970 \\ 
  GRTS-MB & Hg ppb & 50 & 0.2539 & 12.9467 & 0.9215 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 1.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{llrrrr}
  \hline
Approach & Response & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & ZMMI & 50 & -0.1191 & 1.9608 & 0.9355 \\ 
  SRS-MB & ZMMI & 50 & 0.0000 & 2.1477 & 0.9395 \\ 
  GRTS-DB & ZMMI & 50 & -0.0047 & 1.9953 & 0.9210 \\ 
  GRTS-MB & ZMMI & 50 & 0.0461 & 2.4502 & 0.9440 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 2.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{llrrrr}
  \hline
Approach & Response & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Hg ppb & 100 & -0.4156 & 7.0344 & 0.9230 \\ 
  SRS-MB & Hg ppb & 100 & -0.4213 & 7.4820 & 0.9190 \\ 
  GRTS-DB & Hg ppb & 100 & 0.0075 & 7.1415 & 0.9045 \\ 
  GRTS-MB & Hg ppb & 100 & -0.1390 & 8.9667 & 0.9290 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 3.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{llrrrr}
  \hline
Approach & Response & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & ZMMI & 100 & -0.0968 & 1.2958 & 0.9450 \\ 
  SRS-MB & ZMMI & 100 & -0.0942 & 1.3920 & 0.9445 \\ 
  GRTS-DB & ZMMI & 100 & -0.0066 & 1.3219 & 0.9205 \\ 
  GRTS-MB & ZMMI & 100 & -0.0541 & 1.6901 & 0.9460 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 4.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{llrrrr}
  \hline
Approach & Response & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & Hg ppb & 200 & -0.2762 & 4.3998 & 0.9410 \\ 
  SRS-MB & Hg ppb & 200 & -0.1775 & 4.6970 & 0.9400 \\ 
  GRTS-DB & Hg ppb & 200 & -0.0538 & 4.5051 & 0.9325 \\ 
  GRTS-MB & Hg ppb & 200 & -0.0876 & 5.9176 & 0.9450 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 5.} 
\end{table}
\begin{table}[ht]
\centering
\scalebox{0.8}{
\begin{tabular}{llrrrr}
  \hline
Approach & Response & n & MB & RMSE & Coverage \\ 
  \hline
SRS-DB & ZMMI & 200 & -0.0311 & 0.8280 & 0.9450 \\ 
  SRS-MB & ZMMI & 200 & -0.0736 & 0.9090 & 0.9450 \\ 
  GRTS-DB & ZMMI & 200 & 0.0193 & 0.8524 & 0.9460 \\ 
  GRTS-MB & ZMMI & 200 & -0.0093 & 1.1311 & 0.9455 \\ 
   \hline
\end{tabular}
}
\caption{Sampling-analysis combination (Approach), population layout (Layout), response type (Response), proportion of dependent random error (DRE\%), sample size (n), mean bias (MB), root-mean-squared error (rMS(P)E), and 95\% interval coverage (Coverage) in a data application scenario 6.} 
\end{table}

\end{document}