### 2022-04-19 inst/manuscript/manuscript.tex

\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
\usepackage[hyphens]{url}

  \journal{Methods in Ecology and Evolution} % Sets Journal name

\usepackage{lineno} % add
  \linenumbers % turns line numbering on

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\bibliographystyle{elsarticle-harv}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={A comparison of design-based and model-based approaches for finite population spatial sampling and inference.},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{bm} \usepackage{bbm} \usepackage{color} \DeclareMathOperator{\var}{{var}} \DeclareMathOperator{\cov}{{cov}} \usepackage{caption} \usepackage{subcaption} \usepackage{setspace} \usepackage{fancyhdr}

\pagestyle{fancy} \fancyhf{} \chead{Spatial design-based vs model-based}
\doublespacing

\begin{document}

\begin{frontmatter}

  \title{A comparison of design-based and model-based approaches for finite
population spatial sampling and inference.}
    \author[USEPA]{Michael Dumelle\corref{1}}

    \author[STLAW]{Matt Higham}

    \author[NOAA]{Jay M. Ver Hoef}

    \author[USEPA]{Anthony R. Olsen}

    \author[OSU]{Lisa Madsen}

      \address[USEPA]{United States Environmental Protection Agency, 200 SW 35th St,
Corvallis, Oregon, 97333}
    \address[STLAW]{Saint Lawrence University Department of Mathematics, Computer Science,
and Statistics, 23 Romoda Drive, Canton, New York, 13617}
    \address[NOAA]{Marine Mammal Laboratory, Alaska Fisheries Science Center, National
Oceanic and Atmospheric Administration, Seattle, Washington, 98115}
    \address[OSU]{Oregon State University Department of Statistics, 239 Weniger Hall,
Corvallis, Oregon, 97331}
      \cortext[1]{Corresponding Author: Michael Dumelle (Dumelle.Michael@epa.gov)}

  \begin{abstract}
  \begin{enumerate}
  \def\labelenumi{\arabic{enumi}.}
  \tightlist
  \item
    The design-based and model-based approaches to frequentist statistical
    inference rest on fundamentally different foundations. In the
    design-based approach, inference relies on random sampling. In the
    model-based approach, inference relies on distributional assumptions.
    We compare the approaches in a finite population spatial context.
  \item
    We provide relevant background for the design-based and model-based
    approaches and then study their performance using simulated and real
    data from the United States Environmental Protection Agency's 2012
    National Lakes Assessment. A variety of sample sizes, location
    layouts, dependence structures, and response types are considered. The
    population mean is the parameter of interest and performance is
    measured using statistics like bias, squared error, and interval
    coverage.
  \item
    When studying the simulated and real data, we found that regardless of
    the strength of spatial dependence in the data, the generalized random
    tessellation stratified (GRTS) algorithm, which explicitly
    incorporates spatial locations into sampling, tends to outperform the
    simple random sampling (SRS) algorithm, which does not explicitly
    incorporate spatial locations into sampling. We also found that
    model-based inference tends to outperform design-based inference, even
    for skewed data where the model-based distributional assumptions are
    violated. The performance gap between design-based inference and
    model-based inference is small when GRTS samples are used but large
    when SRS samples are used, suggesting that the sampling choice
    (whether to use GRTS or SRS) is most important when performing
    design-based inference.
  \item
    There are many benefits and drawbacks to the design-based and
    model-based approaches for finite population spatial sampling and
    inference that practitioners must consider when choosing between them.
    We provide relevant background contextualizing each approach and study
    their properties in a variety of scenarios, making recommendations for
    use based on the practitioner's goals.
  \end{enumerate}
  \end{abstract}

 \end{frontmatter}

\hypertarget{keywords}{%
\section*{Keywords}\label{keywords}}
\addcontentsline{toc}{section}{Keywords}

Design-based inference; Finite population block kriging (FPBK);
Generalized random tessellation stratified (GRTS) algorithm; Local
neighborhood variance estimator; Model-based inference; Restricted
maximum likelihood (REML) estimation; Spatially balanced sampling;
Spatial covariance

\hypertarget{sec:introduction}{%
\section{Introduction}\label{sec:introduction}}

When data cannot be collected for all units in a population (population
units), data are collected on a subset of the population units -- this
subset is called a sample. There are two general approaches for using
samples to make frequentist statistical inferences about a population:
design-based and model-based. In the design-based approach, inference
relies on randomly assigning some population units to be in the sample
(random sampling). Alternatively, in the model-based approach, inference
relies on distributional assumptions about the underlying
data-generating stochastic process (superpopulation). Each paradigm has
a deep historical context (Sterba, 2009) and its own set of benefits and
drawbacks (Brus and De Gruijter, 1997; Hansen et al., 1983). In this
manuscript, we compare design-based and model-based approaches for
finite population spatial sampling and inference.

Spatial data are data that have some sort of spatial index (usually
specified via coordinates). De Gruijter and Ter Braak (1990) and Brus
and DeGruijter (1993) give early comparisons of design-based and
model-based approaches for spatial data, quashing the belief that
design-based approaches could not be used for spatially correlated data.
Since then, there have been several general comparisons between
design-based and model-based approaches for spatial data (Brus and De
Gruijter, 1997; Brus, 2021; Ver Hoef, 2002, 2008). Cooper (2006) reviews
the two approaches in an ecological context before introducing a
``model-assisted'' variance estimator that combines aspects from each
approach. In addition to Cooper (2006), there has been substantial
research and development into estimators that use both design-based and
model-based principles (see e.g., Sterba (2009) and Cicchitelli and
Montanari (2012), and for Bayesian approaches, see Chan-Golston et al.
(2020) and Hofman and Brus (2021)).

While comparisons between design-based and model-based approaches have
been studied in spatial contexts, our contribution is comparing
design-based approaches specifically built for spatial data to
model-based approaches. Though the broad comparisons we draw between
design-based and model-based approaches generalize to finite and
infinite populations, we focus on finite populations. A finite
population contains a finite number of population units (we assume the
finite number is known) -- an example is lakes (treated as a whole with
the lake centroid representing location) in the conterminous United
States. An infinite population contains an infinite number of population
units -- an example is locations within a single lake.

The rest of the manuscript is organized as follows. In Section
\ref{sec:background}, we introduce and provide relevant background for
design-based and model-based approaches to finite population spatial
sampling and inference. In Section \ref{sec:mm}, we describe how we
intend to compare performance of the approaches using simulated and real
data. The real data is from the United States Environmental Protection
Agency's 2012 National Lakes Assessment (NLA) (USEPA, 2012). In Section
\ref{sec:results}, we present analysis results for the simulated data
and NLA data. And in Section \ref{sec:discussion}, we end with a
discussion and provide directions for future research.

\hypertarget{sec:background}{%
\subsection{Background}\label{sec:background}}

The design-based and model-based approaches incorporate randomness in
fundamentally different ways. In this section, we describe the role of
randomness for each approach and the subsequent effects on statistical
inferences for spatial data.

\hypertarget{subsec:dvm_compare}{%
\subsubsection{Comparing Design-Based and Model-Based
Approaches}\label{subsec:dvm_compare}}

The design-based approach assumes the population is fixed. Randomness is
incorporated via the selection of population units according to a
sampling design. A sampling design assigns a probability of selection to
each sample (subset of population units). Some examples of commonly used
sampling designs include simple random sampling, stratified random
sampling, and cluster sampling. The inclusion probability of a
population unit is calculated by summing each sample's probability of
selection over all samples that contain the population unit. Inclusion
probabilities are often used when selecting samples and estimating
population parameters.

When samples are chosen in a manner such that the layout of sampled
units reflects the layout of the population units, we call the resulting
sample spatially balanced. By ``reflecting the layout of the population
units'', we mean that if population units are concentrated in specific
areas, the units in the sample should be concentrated in the same areas.
Because spatially balanced samples reflect the layout of the population
units, they are not necessarily spread out in space in some equidistant
manner. One method of selecting spatially balanced samples is the
generalized random tessellation stratified (GRTS) algorithm (Stevens and
Olsen, 2004), which we discuss in more detail in Section
\ref{subsec:spb_design}. To quantify the spatial balance of a sample,
Stevens and Olsen (2004) proposed loss metrics based on Voronoi polygons
(i.e., Dirichlet Tessellations).

Fundamentally, the design-based approach combines the randomness of the
sampling design with the data collected via the sample to justify the
estimation and uncertainty quantification of fixed, unknown parameters
of a population (e.g., a population mean). Treating the data as fixed
and incorporating randomness through the sampling design yields
estimators having very few other assumptions. Confidence intervals for
these types of estimators are typically derived using limiting arguments
that incorporate all possible samples. Sample means, for example, are
asymptotically normal (Gaussian) by the central limit theorem (under
some assumptions). If we repeatedly select samples from the population,
then 95\% of all 95\% confidence intervals constructed from a procedure
with appropriate coverage will contain the true fixed population mean.
Särndal et al. (2003) and Lohr (2009) provide thorough reviews of the
design-based approach.

The model-based approach assumes the population is a random realization
of a data-generating stochastic process. Randomness is formally
incorporated through distributional assumptions on this process.
Strictly speaking, randomness need not be incorporated through random
sampling, though Diggle et al. (2010) warn against preferential
sampling. Preferential sampling occurs when the process generating the
data locations and the process being modeled are not independent of one
another. To guard against preferential sampling, model-based approaches
can implement some form of random sampling. It is common, however, for
model-based approaches to sample non-randomly. When model-based
approaches do implement random sampling, the inclusion probabilities are
ignored when analyzing the sample (in contrast to the design-based
approach, which relies on these inclusion probabilities to analyze the
sample).

Instead of estimating fixed, unknown population parameters, as in the
design-based approach, often the goal of model-based inference is to
predict the value of a realized variable. For example, suppose the
realized mean of all population units (the realized population mean) is
the variable of interest. Instead of a fixed, unknown mean, we are
predicting the value of the mean, a random variable. Prediction
intervals are then derived using assumptions of the data-generating
stochastic process. If we repeatedly generate realizations from the same
process and select samples, then 95\% of all 95\% prediction intervals
constructed from a procedure with appropriate coverage will contain
their respective realized means. Cressie (1993) and Schabenberger and
Gotway (2017) provide thorough reviews of model-based approaches for
spatial data. In Fig. \ref{fig:fig1}, we provide a visual comparison of
the design-based and model-based approaches (Ver Hoef (2002) and Brus
(2021) provide similar figures). Fig. \ref{fig:fig1} contrasts the
design-based approach with a fixed population and random sampling to the
model-based approach with random populations and non-random sampling.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/dvm_comp.jpeg}
  \caption{A visual comparison of the design-based and model-based approaches. In the top row, the design-based approach is highlighted. There is one fixed population with nine population units and three random samples of size four (points circled are those sampled). The response values at each site are fixed. In the bottom row, the model-based approach is highlighted. There are three realizations of the same data-generating stochastic process that are all sampled at the same four locations. The response values at each site are random.}
  \label{fig:fig1}
\end{figure}

\hypertarget{subsec:spb_design}{%
\subsubsection{Spatially Balanced Design and
Analysis}\label{subsec:spb_design}}

We previously mentioned that the design-based approach can be used to
select spatially balanced samples. Spatially balanced samples are useful
because parameter estimates from these samples tend to vary less than
parameter estimates from samples lacking spatial balance (Barabesi and
Franceschi, 2011; Benedetti et al., 2017; Grafström and Lundström, 2013;
Robertson et al., 2013; Stevens and Olsen, 2004; Wang et al., 2013). The
first spatially balanced sampling algorithm to see widespread use was
the generalized random tessellation stratified (GRTS) algorithm (Stevens
and Olsen, 2004). After the GRTS algorithm was developed, several other
spatially balanced sampling algorithms emerged, including stratified
sampling with compact geographical strata (Walvoort et al., 2010), the
local pivotal method (Grafström et al., 2012; Grafström and Matei,
2018), spatially correlated Poisson sampling (Grafström, 2012), balanced
acceptance sampling (Robertson et al., 2013), within-sample-distance
sampling (Benedetti and Piersimoni, 2017), and Halton iterative
partitioning sampling (Robertson et al., 2018). In this manuscript, we
select spatially balanced samples using the GRTS algorithm because it is
readily available in the \texttt{spsurvey} \textbf{\textsf{R}} package
(Dumelle et al., 2022) and naturally accommodates finite and infinite
sampling frames, unequal inclusion probabilities, and replacement units.
Replacement units are additional population units that can be sampled
when a population unit originally selected can no longer be sampled. A
couple of reasons why an originally selected site can no longer be
sampled include its location being physically inaccessible or it is on
private land that the researcher does not have permission to access.

The GRTS algorithm selects samples by utilizing a particular mapping
between two-dimensional and one-dimensional space that preserves
proximity relationships. First, the bounding box of the domain is split
up into four distinct, equally sized squares called level-one cells.
Each level-one cell is randomly assigned a level-one address of 0, 1, 2,
or 3. The set of level-one cells is denoted by \(\mathcal{A}_1\) and
defined as \(\mathcal{A}_1 \equiv \{a_1: a_1 = 0, 1, 2, 3\}\). Within
each level-one cell, the inclusion probability for each population unit
(which is pre-specified) is summed, and if any of these sums exceed one,
a second level of cells is added. Then each level-one cell is split into
four distinct, equally sized squares called level-two cells. Each
level-two cell is randomly assigned a level-two address of 0, 1, 2, or
3. The set of level-two cells is denoted by \(\mathcal{A}_2\) and
defined as
\(\mathcal{A}_2 \equiv \{a_1a_2: a_1 = 0, 1, 2, 3; a_2 = 0, 1, 2, 3\}\).
The inclusion probabilities within each level-two cell are summed, and
if any of these sums exceed one, a third level of cells is added. This
process continues for \(k\) steps, until all level-\(k\) cells have
inclusion probability sums no larger than one. Then
\(\mathcal{A}_k \equiv \{a_1...a_k : a_1 = 0, 1, 2, 3; ...; a_k = 0, 1, 2, 3\}\).
Figure\(~\)\ref{fig:grts} provides some intuition regarding the
assignment of level-one and level-two cells.

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/grts_level1.jpeg}
  \caption{Assignment of level-one cells to the spatial domain. Grey circles indicate population units.}
  \label{fig:grts_level1}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/grts_level2.jpeg}
  \caption{Assignment of level-two cells to the spatial domain. Grey circles indicate population units.}
  \label{fig:grts_level2}
\end{subfigure}
\caption{Assignment of level-one and level-two cells to the spatial domain. In (a), each level-one cells is randomly given a level-one address of 0, 1, 2, or 3. In (b), each level-two cell within each level-one cell is randomly given a level-two address of 0, 1, 2, or 3.}
\label{fig:grts}
\end{figure}

After determining \(\mathcal{A}_k\), the set is placed into hierarchical
order. Hierarchical order is a numeric order that first sorts
\(\mathcal{A}_k\) by the level-one addresses from smallest to largest,
then sorts \(\mathcal{A}_k\) by the level-two addresses from smallest to
largest, and so on. For example, \(\mathcal{A}_2\) in hierarchical order
is the set
\linebreak \(\{00, 01, 02, 03, 10, ..., 13, 20, ..., 23, 30, ..., 33\}\).
Because hierarchical ordering sorts by level-one cells, then level-two
cells, and so on, population units that have similar hierarchical
addresses tend to be nearby one another in space. Next, each population
unit is mapped to a one-dimensional line in hierarchical order where
each population unit's inclusion probability equals its line-length. If
a level-\(k\) cell has multiple population units in it, they are
randomly placed within the cell's respective line segment. A uniform
random variable is then simulated in \([0, 1]\) and a systematic sample
is selected on the line, yielding \(n\) sample points for a sample size
\(n\). Each of these sample points falls on some population unit's line
segment, and thus that population unit is selected in the sample. For
further details regarding the GRTS algorithm, see Stevens and Olsen
(2004).

After selecting a sample and collecting data, unbiased estimates of
population means and totals can be obtained using the Horvitz-Thompson
estimator (Horvitz and Thompson, 1952). If \(\tau\) is a population
total, the Horvitz-Thompson estimator for \(\tau\), denoted by
\(\hat{\tau}_{ht}\), is given by \begin{align}\label{eq:ht}
  \hat{\tau}_{ht} = \sum_{i = 1}^n z_i \pi_i^{-1},
\end{align} where \(z_i\) is the value of the \(i\)th population unit in
the sample, \(\pi_i\) is the inclusion probability of the \(i\)th
population unit in the sample, and \(n\) is the sample size. An estimate
of the population mean is obtained by dividing \(\hat{\tau}_{ht}\) by
\(N\), the number of population units.

It is also important to quantify the uncertainty in \(\hat{\tau}_{ht}\).
The Horvitz-Thompson (Horvitz and Thompson, 1952) and Sen-Yates-Grundy
(Sen, 1953; Yates and Grundy, 1953) variance estimators are often used
to estimate \(\text{Var}(\hat{\tau}_{ht})\), but these estimators have
two drawbacks. First, they rely on calculating \(\pi_{ij}\), the
probability that population unit \(i\) and population unit \(j\) are
both in the sample -- this quantity can be challenging if not impossible
to calculate analytically for GRTS samples. Second, these estimators
tend to ignore the spatial locations of the population units. To address
these two drawbacks simultaneously, Stevens and Olsen (2003) proposed
the local neighborhood variance estimator. The local neighborhood
variance estimator does not rely on \(\pi_{ij}\) and estimates the
variance of \(\hat{\tau}\) conditional on the random properties of the
GRTS sample -- the idea being that this conditioning should yield a more
precise estimate of \(\hat{\tau}\). They show that the contribution from
each sampled population unit to the overall variance is dominated by
local variation. Thus the local neighborhood variance estimator is a
weighted sum of variance estimates from each sampled population unit's
local neighborhood. These local neighborhoods contain the sampled
population unit itself and its three nearest neighbors (among all other
sampled population units). For more details, see Stevens and Olsen
(2003).

\hypertarget{finite-population-block-kriging}{%
\subsubsection{Finite Population Block
Kriging}\label{finite-population-block-kriging}}

Finite population block kriging (FPBK) is a model-based approach that
expands the geostatistical Kriging framework to the finite population
setting (Ver Hoef, 2008). Instead of developing inference based on a
specific sampling design, we assume the data are generated by a spatial
stochastic process. We summarize some of the basic principles of FPBK
next -- see Ver Hoef (2008) for technical details and see Higham, J. Ver
Hoef, et al. (2021) for an extension to cases of imperfect detection
among population units. Let
\({\mathbf{z} \equiv \{\text{z}(s_1), \text{z}(s_2), . . . , \text{z}(s_N) \}}\)
be an \(N \times 1\) response vector at locations \(s_1\), \(s_2\), . .
. , \(s_N\) that can be measured at the \(N\) population units. Suppose
we want to use a sample to predict some linear function of the response
variable, \(f(\mathbf{z}) = \mathbf{b}^\prime \mathbf{z}\), where
\(\mathbf{b}^\prime\) is a \(1 \times N\) vector of weights (e.g, the
population mean is represented by a weights vector whose elements all
equal \(1 / N\)). Denoting quantities that are part of the sampled
population units with a subscript \emph{s} and quantities that are part
of the unsampled population units with a subscript \emph{u}, let

\begin{equation}
\begin{pmatrix} \label{equation:Zmarginal}
    \mathbf{z}_s      \\
    \mathbf{z}_u
\end{pmatrix}
=
\begin{pmatrix}
  \mathbf{X}_s    \\
  \mathbf{X}_u
\end{pmatrix}
\bm{\beta} +
\begin{pmatrix}
\bm{\delta}_s    \\
\bm{\delta}_u
\end{pmatrix},
\end{equation} where \(\mathbf{X}_s\) and \(\mathbf{X}_u\) are the
design matrices for the sampled and unsampled population units,
respectively, \(\bm{\beta}\) is the parameter vector of fixed effects,
and \(\bm{\delta} \equiv [\bm{\delta}_s \,\, \bm{\delta}_u]'\), where
\(\bm{\delta}_s\) and \(\bm{\delta}_u\) are random errors for the
sampled and unsampled population units, respectively.

FPBK assumes \(\bm{\delta}\) in Equation\(~\)\eqref{equation:Zmarginal}
has mean-zero and a spatial dependence structure that can be modeled
using a covariance function. This covariance function is commonly
assumed to be non-negative, second-order stationary (depending only on
the separation vector (e.g., distance) between population units),
isotropic (independent of direction), and decays with distance between
population units (Cressie, 1993). Henceforth, it is implied that we have
made these same assumptions regarding \(\bm{\delta}\). Chiles and
Delfiner (1999), pp.~80-93 discuss covariance functions that are not
second-order stationary, not isotropic, or not either. A variety of
flexible covariance functions can be used to model \(\bm{\delta}\)
(Cressie, 1993) -- one example is the exponential covariance function.
Cressie (1993) provides a thorough list of spatial covariance functions.
The \(i,j\)th element of the exponential covariance matrix,
\(\mathop{\mathrm{{cov}}}(\bm{\delta})\), is \mbox{}
\begin{align}\label{equation:expcov}
\mathop{\mathrm{{cov}}}(\delta_i, \delta_j) = 
\begin{cases} 
\sigma^2_{1}\exp(-h_{i,j}/\phi) & h_{i,j} > 0 \\
\sigma^2_{1} + \sigma^2_2 & h_{i,j} = 0
\end{cases}
,
\end{align} where \(\sigma^2_{1}\) is the variance parameter that
quantifies the spatially dependent (correlated) variability,
\(\sigma^2_{2}\) is the variance parameter the quantifies that spatially
independent (not correlated) variability, \(\phi\) is the distance
parameter that measures the distance-decay rate of the covariance, and
\(h_{i,j}\) is the Euclidean distance between population units \(i\) and
\(j\). In geostatistical literature, \(\sigma^2_{1}\) is called the
partial sill, \(\sigma^2_{2}\) is called the nugget, and \(\phi\) is
called the range. We denote \(\bm{\theta}\) as the vector of covariance
parameters that composes \(\bm{\delta}\). In
Equation\(~\)\ref{equation:expcov},
\(\bm{\theta} = \{\sigma^2_1, \sigma^2_2, \phi \}\).

The parameters in Equation\(~\)\ref{equation:Zmarginal} can be estimated
using a variety of techniques, but we focus on using restricted maximum
likelihood (REML) (Harville, 1977; Patterson and Thompson, 1971;
Wolfinger et al., 1994). REML is preferred over maximum likelihood (ML)
because ML estimates can be badly biased for small sample sizes, due to
the fact that ML makes no adjustment for the simultaneous estimation of
\(\bm{\beta}\) and \(\bm{\theta}\) (Patterson and Thompson, 1971). Minus
twice the REML log-likelihood of the sampled sites is given by
\begin{equation}\label{equation:reml}
  \ln|\bm{\Sigma}| + (\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}})^T \bm{\Sigma}_{ss}^{-1}(\bm{z}_s - \bm{X}_s \bm{\tilde{\beta}}) + \ln|\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s| + (n - p) \ln(2 \pi) ,
\end{equation} where
\(\bm{\tilde{\beta}} = (\bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{X}_s)^{-1} \bm{X}_s^T \bm{\Sigma}_{ss}^{-1} \bm{z}_s\)
and \(\bm{\Sigma}_{ss}\) is the covariance matrix of the sampled sites.
Minimizing Equation\(~\)\ref{equation:reml} yields
\(\bm{\hat{\theta}}_{reml}\), the REML estimates of \(\bm{\theta}\).
Then \(\bm{\hat{\beta}}_{reml}\), the REML estimate of \(\bm{\beta}\),
is given by
\((\bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{X})^{-1} \bm{X}_s^T \bm{\hat{\Sigma}}_{ss}^{-1} \bm{z}_s\),
where \(\bm{\hat{\Sigma}}_{ss}\) is \(\bm{\Sigma}_{ss}\) evaluated at
\(\bm{\hat{\theta}}_{reml}\).

With the model formulation in Equation\(~\)\ref{equation:Zmarginal}, the
best linear unbiased predictor (BLUP) of \(f(\mathbf{b}'\mathbf{z})\)
and its prediction variance can be computed. While details of the
derivation are in Ver Hoef (2008), we note here that the predictor and
its variance are both moment-based, meaning that they do not rely on any
distributional assumptions. Distributional assumptions are used,
however, when constructing prediction intervals.

Other approaches, such as k-nearest-neighbors (Fix and Hodges, 1989; Ver
Hoef and Temesgen, 2013) and random forest (Breiman, 2001), among
others, could also be used to obtain predictions for a mean or total
from finite population spatial data. Compared to the k-nearest-neighbors
and random forest approach, we prefer FPBK because it is model-based and
relies on theoretically-based variance estimators leveraging the model's
spatial covariance structure, whereas k-nearest-neighbors and random
forests use ad-hoc variance estimators (Ver Hoef and Temesgen, 2013).
Additionally, Ver Hoef and Temesgen (2013) compared FPBK,
k-nearest-neighbors, and random forest in a variety of spatial data
contexts, and FPBK tended to perform best.

\hypertarget{sec:mm}{%
\section{Materials and Methods}\label{sec:mm}}

In this section we describe how we used simulated and real data to
investigate performance between simple random sampling (SRS) and GRTS
sampling as well as performance between design-based (DB) and
model-based (MB) inference. In SRS and GRTS sampling, all population
units had equal inclusion probabilities and were selected without
replacement. The important distinction between SRS and GRTS is that SRS
ignores spatial locations while sampling but GRTS explicitly
incorporates them. Together, the two sampling plans (SRS and GRTS)
combined with the two inference approaches (DB and MB) yielded four
sampling-inference combinations: SRS-DB, SRS-MB, GRTS-DB, and GRTS-MB.
For SRS-DB, the Horvitz-Thompson estimator \eqref{eq:ht} was used to
estimate means and the commonly-used SRS variance formula (Lohr, 2009;
Särndal et al., 2003) was used to estimate the variance. This variance
formula is given by \begin{equation}\label{equation:srs_var}
 \frac{f[\sum_{i = 1}^n (z_i - \bar{z})^2]}{n(n - 1)},
\end{equation} where \(z_i\) is the \(i\)th response value, \(\bar{z}\)
is the mean of all \(z_i\), \(n\) is the sample size, \(N\) is the
population size, and \(f = (1 - n / N)\) (\(f\) is often called the
finite population correction factor). For GRTS-DB, the Horvitz-Thompson
estimator was used to estimate means and the local neighborhood variance
was used to estimate variances. For SRS-MB and GRTS-MB, FPBK was used to
estimate means and variances using restricted maximum likelihood. SRS,
GRTS sampling, and design-based inference were implemented using the
\texttt{spsurvey} \textbf{\textsf{R}} package (Dumelle et al., 2022).
FPBK was implemented using the \texttt{sptotal} \textbf{\textsf{R}}
package (Higham, Ver Hoef, et al., 2021).

We used simulated data to compare the sampling-inference combinations
across many realized populations from the same data-generating
stochastic process. We then used real data from the 2012 National Lakes
Assessment to compare the sampling-inference combinations within a
single realized population (which is typically the case in reality).
With the simulated data, we were in control of the data-generating
stochastic process and the random sampling process. With the real data,
we were in control of only the random sampling process.

\hypertarget{sec:mm_sim}{%
\subsection{Simulated Data}\label{sec:mm_sim}}

We evaluated performance of the four sampling-inference combinations in
36 different simulation scenarios. The 36 scenarios resulted from the
crossing of three sample sizes, two location layouts (of the population
units), two response types, and three proportions of dependent random
error (DRE). The three sample sizes (\(n\)) were \(n = 50, n = 100,\)
and \(n = 200\). Samples were always selected from a population size
(\(N\)) of \(N = 900\). The two location layouts were random and
gridded. Locations in the random layout were randomly generated inside
the unit square (\([0, 1] \times [0, 1]\)). Locations in the gridded
layout were placed on a fixed, equally spaced grid inside the unit
square. The two response types were normal and skewed. For the normal
response type, the response was simulated using mean-zero random errors
with the exponential covariance (Equation\(~\)\ref{equation:expcov}) for
three proportions of dependent random error (DRE): 0\% DRE, 50\% DRE,
and 90\% DRE. Recall the proportion of DRE is represented by
\(\sigma^2_1 / (\sigma^2_1 + \sigma^2_2)\), where \(\sigma^2_1\) and
\(\sigma^2_2\) are the DRE variance and independent random error (IRE)
variance from Equation\(~\)\ref{equation:expcov}, respectively. The
total variance, \(\sigma^2_1 + \sigma^2_2\), was always 2. The distance
parameter was always \(\sqrt{2} / 3\), chosen so that the correlation in
the DRE decayed to nearly zero at \(\sqrt{2}\), the largest possible
distance between two population units in the domain. For the skewed
response type, the response was first simulated using the same approach
as for the normal response type, except that the total variance was
0.6931 instead of 2. The response was then exponentiated, yielding a
skewed random variable whose total variance was 2. The skewed responses
were used to evaluate performance of the sampling-inference approaches
for data that were not normal but were still estimated using REML, which
relies on a normal log-likelihood. Figure \ref{fig:sim_pops} shows an
example of a realized population for the normal and skewed responses
using the random layout and 50\% DRE.

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/symm_pop_hist.jpeg}
  \caption{Histogram of a realized population for the normal response.}
  \label{fig:symm_pop_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/skew_pop_hist.jpeg}
  \caption{Histogram of a realized population for the skewed response.}
  \label{fig:skew_pop_hist}
\end{subfigure}
\caption{Histograms of realized populations simulated for the normal and skewed resposnes using the random layout and 50\% DRE.}
\label{fig:sim_pops}
\end{figure}

In each of the 36 simulation scenarios, there were 2000 independent
simulation trials. Within each simulation scenario and trial, IRS and
GRTS samples were selected and then design-based and model-based
inferences were used to estimate (design-based) or predict (model-based)
the mean and construct 95\% confidence (design-based) or 95\% prediction
(model-based) intervals. With model-based inference, covariance
parameters and the mean were estimated (using REML) separately for each
trial. After all 2000 trials, we summarized the long-run performance of
the sampling-inference combination in each scenario by calculating mean
bias, root-mean-squared error, and interval coverage. Mean bias is taken
as the average deviation between each trial's estimated (or predicted)
mean (\(\hat{\mu}_i\)) and its realized mean (\(\mu_i\)):
\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)\), where \(i\)
indexes simulation trials. Root-mean-squared error is taken as the
square root of the average squared deviation between each trial's
estimated (or predicted) mean and its realized mean:
\(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu_i)^2}\).
Interval coverage is taken as the proportion of simulation trials where
the realized mean is contained in its 95\% confidence (or prediction)
interval. These intervals are constructed using the normal distribution
-- justification comes from the asymptotic normality of means via the
central limit theorem (under some assumptions). Quantifying these
metrics is important because together, they give us an idea of the
accuracy (mean bias), spread (RMSE), and validity (interval coverage) of
the sampling-inference combinations.

\hypertarget{sec:mm_app}{%
\subsection{National Lakes Assessment Data}\label{sec:mm_app}}

The United States Environmental Protection Agency (USEPA), states, and
tribes periodically conduct National Aquatic Research Surveys (NARS) to
assess the water quality of various bodies of water in the conterminous
United States. One component of NARS is the National Lakes Assessment
(NLA), which measures various aspects of lake health and water quality.
We focus on analyzing zooplankton multi-metric indices (ZMMI) and
mercury concentrations in parts per billion (Hg ppb) from the 2012 NLA.
For ZMMI, data were collected at 1035 unique lakes. At less than 10\% of
lakes, two ZMMI replicates were collected. These were averaged for the
purposes of our study so that each lake had one measurement for ZMMI.
For Hg ppb, data were collected at 995 unique lakes (there were no
replicates). The ZMMI and Hg ppb data are shown as spatial maps and as
histograms in Figure \ref{fig:zmmi}. The ZMMI data tend to be highest
near the coasts, lowest in the Central United States, are relatively
symmetric, and have a mean of 55.05. The Hg ppb data tend to be highest
in the Northeastern United States, lowest elsewhere, are skewed, and
have a mean of 103.16 ppb. Also in Figure \ref{fig:zmmi} are separate
spatial semivariograms for ZMMI and Hg ppb. The spatial semivariogram
quantifies the halved average squared differences (semivariance) of
responses whose separation (distance) falls within some distance class.
The spatial semivariance is closely related to the spatial covariance,
and spatial semivariograms are often used to gauge the strength of
spatial dependence in data. Both ZMMI and Hg ppb seem to have moderately
strong spatial dependence (Figure \ref{fig:zmmi}), as the semivariance
increases steadily with distance (meaning that observations nearby one
another tend to be more similar than observations far apart from one
another).

\begin{figure}
\centering
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/zmmi_map.jpeg}
  \caption{Spatial map of the ZMMI population.}
  \label{fig:zmmi_map}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/mercury_map.jpeg}
  \caption{Spatial map of the Hg ppb population.}
  \label{fig:mercury_map}
\end{subfigure} \\
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/zmmi_hist.jpeg}
  \caption{Histogram of the ZMMI population.}
  \label{fig:zmmi_hist}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/mercury_hist.jpeg}
  \caption{Histogram of the Hg ppb population.}
  \label{fig:mercury_hist}
\end{subfigure} \\
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/zmmi_sv.jpeg}
  \caption{Semivariogram of the ZMMI population.}
  \label{fig:zmmi_sv_plot}
\end{subfigure}
\begin{subfigure}{0.49\textwidth}
  \centering
  \includegraphics[width = 1\linewidth]{figures/mercury_sv.jpeg}
  \caption{Semivariogram of the Hg ppb population.}
  \label{fig:mercury_sv_plot}
\end{subfigure}
\caption{Exploratory graphics representing populations for the zooplankton multi-metric indices (ZMMI) and mercury concentration in parts per billion (Hg ppb) in the 2012 National Lakes Assessment (NLA) data.}
\label{fig:zmmi}
\end{figure}

We studied performance of the four sampling-inference combinations by
selecting 2000 random IRS and GRTS samples of size \(n = 50\),
\(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
populations and then analyzing the samples using MB and DB inference. In
total, there were six separate scenarios (two responses crossed with
three sample sizes). For each sample from each scenario, parameters were
estimated using the same approaches as described for simulated data. We
also used the same evaluation metrics as for the simulated data: mean
bias, RMSE, and interval coverage. Mean bias is taken as the average
deviation between each sample's estimated (or predicted) mean
(\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of ZMMI or Hg ppb):
\(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\), where \(i\)
indexes simulation trials. Root-mean-squared error is taken as the
square root of the average squared deviation between each sample's
estimated (or predicted) mean and its population mean:
\(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
Interval coverage is taken as the proportion of simulation trials where
the population mean is contained in its 95\% confidence (or prediction)
interval. These intervals are constructed using the normal distribution.

\hypertarget{sec:results}{%
\section{Results}\label{sec:results}}

\hypertarget{sec:r_sim}{%
\subsection{Simulated Data}\label{sec:r_sim}}

Mean bias is nearly zero for all four sampling-inference combinations in
all 36 scenarios, so we omit a more detailed summary of those results
here. Tables for mean bias in all 36 simulation scenarios are provided
in the supporting information.

We define the relative RMSE as a ratio with numerator given by the RMSE
for a sampling-inference combination and the denominator given by the
RMSE for SRS-DB. Relative RMSEs for the random location layout are
provided in Fig. \ref{fig:rmspe_eff}. When there is no spatial
covariance (Fig. \ref{fig:rmspe_eff}, ``DRE\%: 0\%''), the four
sampling-inference combinations have approximately equal RMSE. In these
scenarios, using GRTS sampling or model-based inference does not
generally increase efficiency compared to SRS-DB. When there is spatial
covariance (Fig. \ref{fig:rmspe_eff}, ``DRE\%: 50\%'' and ``DRE\%:
90\%''), GRTS-MB tends to have the lowest RMSE, followed by GRTS-DB,
SRS-MB, and finally SRS-DB. As the strength of spatial covariance
increases, the gap in RMSE between SRS-DB and the other
sampling-inference combinations widens. Finally we note that when there
is spatial covariance, SRS-MB has a much lower RMSE than SRS-DB,
suggesting that the lack of efficiency from SRS is largely mitigated by
model-based inference. These RMSE conclusions are similar to those
observed in the grid location layout, so we omit a figure and discussion
regarding the grid location layout here. Tables for RMSE in all 36
simulation scenarios are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/rmspe_eff.jpeg}
  \caption{Simulated data relative RMSE for the four sampling-inference combinations and three sample sizes in the random location layout. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black lines separate the sample sizes.}
  \label{fig:rmspe_eff}
\end{figure}

95\% interval coverage for each of the four sampling-inference
combinations in the random location layout is shown in Fig.
\ref{fig:figconf}. Within each simulation scenario, all
sampling-inference combinations tend to have fairly similar interval
coverage, though when \(n = 50\) or \(n = 100\), GRTS-DB coverage is
usually a few percentage points lower than the other combinations, which
suggests that the local neighborhood variance estimate may be slightly
too small for small \(n\). Coverage in the normal response scenarios is
usually near 95\%, while coverage in the skewed response scenarios
usually varies from 90\% to 95\% but increases with the sample size. At
a sample size of 200, all four sampling-inference combinations have
approximately 95\% interval coverage in both response scenarios for all
DRE proportions. These interval coverage conclusions are similar to
those observed in the grid location layout, so we omit a figure and
discussion regarding the grid location layout here. Tables for interval
coverage in all 36 simulation scenarios are provided in the supporting
information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/coverage.jpeg}
  \caption{Simulated data interval coverage for the four sampling-inference combinations and three sample sizes in the random location layout. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
  \label{fig:figconf}
\end{figure}

\hypertarget{sec:r_app}{%
\subsection{National Lakes Assessment Data}\label{sec:r_app}}

Mean bias is nearly zero for all four sampling-inference combinations in
all six scenarios, so we omit a more detailed summary of those results
here. Tables for mean bias in all six simulation scenarios are provided
in the supporting information.

The relative RMSE of both ZMMI (symmetric response) and Hg ppb (skewed
response) for all four sampling-inference combinations are shown in
Fig.\(~\)\ref{fig:data_rmspe_eff}. GRTS-MB has the lowest RMSE, followed
by GRTS-DB, SRS-MB, and then SRS-DB. The difference in RMSE among
GRTS-MB and GRTS-DB tends to be quite small. When \(n = 50\), SRS-MB
RMSE is approximately evenly between IRS-DB RMSE and GRTS-MB RMSE, but
for the larger sample sizes (\(n = 100\), \(n = 200\)), SRS-MB RMSE is
closer to GRTS-MB RMSE. Lastly we note that GRTS-MB, GRTS-DB, and SRS-MB
all have noticeably lower RMSE than SRS-DB. Tables for RMSE in all six
scenarios are provided in the supporting information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/data_rmspe_eff.jpeg}
  \caption{NLA data relative RMSE for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid, black lines separate the sample sizes.}
  \label{fig:data_rmspe_eff}
\end{figure}

95\% interval coverage of both ZMMI and Hg ppb for all four
sampling-inference combinations is shown in Fig. \ref{fig:data_figconf}.
When \(n = 50\), interval coverage for both responses is too low, though
interval coverage is higher for ZMMI (symmetric response) than for Hg
ppb (skewed response). When \(n = 100\), ZMMI interval coverage is
approximately 95\% except for GRTS-DB, which has coverage around 92\%,
while Hg ppb interval coverage ranges from approximately 90\% (GRTS-DB)
to 93\% (GRTS-MB). When \(n = 200\), ZMMI interval coverage is
approximately 95\% while Hg ppb interval coverage ranges from
approximately 93\% (GRTS-DB) to 95\% (GRTS-MB). As with the simulated
data, coverages for the NLA data tend to increase with the sample sizes,
coverages tend to be higher for symmetric responses than for skewed
responses, and the local neighborhood variance was slightly too small
for small \(n\), yielding slightly lower interval coverages than the
other sampling-inference combinations. Recall that model-based inference
defines interval coverage properties across realized populations. With
the simulated data, we evaluated interval coverage across realized
populations, but for the NLA data, we evaluated interval coverage within
a single realized population for different samples. We did find that
model-based coverages were similar to the design-based coverages,
however, suggesting that for some realized populations it is reasonable
to heuristically view data from separate random samples as being from
approximately separate realized populations. But generally, if
model-based intervals constructed from many random samples of a single
realized population show improper coverage, this does not necessarily
imply a deficiency in model-based inference. Tables for interval
coverage in all six simulation scenarios are provided in the supporting
information.

\begin{figure}
  \centering
  \includegraphics[width = 1\linewidth]{figures/data_coverage.jpeg}
  \caption{NLA data interval coverage for the four sampling-inference combinations. The rows indicate the proportion of dependent error and the columns indicate the response type. The solid black lines separate the sample sizes and the dashed black lines represent 95\% coverage.}
  \label{fig:data_figconf}
\end{figure}

\hypertarget{sec:discussion}{%
\section{Discussion}\label{sec:discussion}}

The design-based and model-based approaches to frequentist statistical
inference rest on fundamentally different foundations. Design-based
approaches rely on random sampling to estimate population parameters.
Model-based approaches rely on distributional assumptions to predict
realized values of a data-generating stochastic process. Though
model-based approaches do not rely on random sampling, it can still be
beneficial as a way to guard against preferential sampling. While
design-based and model-based approaches have often been compared in the
literature from theoretical and analytical perspectives, our
contribution lies in studying them for finite population spatial data
while implementing GRTS sampling and the local neighborhood variance
estimator. Aside from the theoretical differences described throughout
the manuscript, a few analytical findings from the simulated and real
data studies were particularly notable. All sampling-inference
combinations had approximately zero mean bias. Independent of the
inference approach, GRTS-DB and GRTS-MB had lower RMSE than their SRS
counterparts. Though GRTS-DB and GRTS-MB generally had very similar
RMSE, SRS-MB tended to have much lower RMSE than SRS-DB, suggesting that
the model-based inference mitigated much of the inefficiency in RMSE
from SRS. As the proportion of dependent random error in the simulated
data increased, SRS-MB, GRTS-DB, and GRTS-MB become increasingly more
efficient (lower RMSE) than SRS-DB. Interval coverage tended to be
higher for the symmetric responses than skewed responses and tended to
increase with the sample size. At a sample size of \(n = 200\),
generally all interval coverages were near the desired value of 95\%.

There are several benefits and drawbacks of the design-based and
model-based approaches for finite population spatial sampling and
inference. Some we have discussed, but others we have not, and they are
worthy of consideration in future research. First, we discuss advantages
of the design-based approach. Design-based inference is often
computationally efficient, while model-based inference can be
computationally burdensome, especially for likelihood-based estimation
methods like REML that rely on the inverse of a covariance matrix.
Design-based inference easily handles binary data through a
straightforward application of the Horvitz-Thompson estimator. In
contrast, analyzing binary data using model-based inference generally
requires a logistic mixed regression model, the parameters of which can
be difficult to estimate and interpret (Bolker et al., 2009). An
advantage of design-based inference is that interval coverage is valid
(has the proper coverage rate) as long as 1) the sample is sufficiently
large to ensure the statistic's sampling distribution is approximately
normal and 2) the variance estimator is consistent (Brus and De
Gruijter, 1997; Särndal et al., 2003). This is because with the
design-based approach, the sampling plan and inclusion probabilities are
specified directly by the researcher. An advantage of SRS-DB not
previously mentioned is that it is likely to be valid given the
consistency of its variance estimator (Särndal et al., 2003). With the
model-based approach, however, interval coverage is unlikely to be valid
if the model assumptions made do not not accurately reflect reality.
Whether model assumptions accurately reflect reality can be a
challenging and sometimes impossible question to answer definitively.

Now, we discuss advantages of the model-based approach. The model-based
approach can more naturally quantify the relationship between covariates
(predictor variables) and the response variable than design-based
approaches. Model-based inference also yields estimated spatial
covariance parameters, which help better understand the dependence
structure of the process in study. Model selection is also possible
using model-based inference and criteria such as cross validation,
likelihood ratio tests, or AIC (Akaike, 1974). Model-based inference is
capable of more efficient small-area estimation than design-based
inference because model-based inference can leverage distributional
assumptions in areas with few observed population units. Model-based
approaches also accommodate unit-by-unit predictions at unobserved
locations that can be used to construct informative visualizations like
smoothed maps. Brus and De Gruijter (1997) provide a more thorough
discussion regarding the benefits and drawbacks of the two approaches.
In short, when deciding whether the design-based or model-based approach
is more appropriate to implement, the benefits and drawbacks of each
approach should be considered alongside the particular goals of the
study.

There are many extensions of this research worthy of future
consideration that include sampling with unequal inclusion
probabilities, using different spatially balanced sampling approaches
(instead of GRTS), using different spatial data configurations, using
different spatial domains like stream networks (Ver Hoef and Peterson,
2010), using different response or covariance structures, and using
spatial or external mean trends (which can be defined through
covariates).

\hypertarget{acknowledgments}{%
\section*{Acknowledgments}\label{acknowledgments}}
\addcontentsline{toc}{section}{Acknowledgments}

We would like to thank the editors and anonymous reviewers for hard work
and time spent providing us with thoughtful, valuable feedback which
greatly improved the manuscript.

The views expressed in this manuscript are those of the authors and do
not necessarily represent the views or policies of the U.S.
Environmental Protection Agency or the National Oceanic and Atmospheric
Administration. Any mention of trade names, products, or services does
not imply an endorsement by the U.S. government, the U.S. Environmental
Protection Agency, or the National Oceanic and Atmospheric
Administration. The U.S. Environmental Protection Agency and National
Oceanic and Atmospheric Administration do not endorse any commercial
products, services, or enterprises.

\hypertarget{conflict-of-interest-statement}{%
\section*{Conflict of Interest
Statement}\label{conflict-of-interest-statement}}
\addcontentsline{toc}{section}{Conflict of Interest Statement}

There are no conflicts of interest for any of the authors.

\hypertarget{author-contribution-statement}{%
\section*{Author Contribution
Statement}\label{author-contribution-statement}}
\addcontentsline{toc}{section}{Author Contribution Statement}

All authors conceived the ideas; All authors designed the methodology;
MD and MH performed the simulations and analyzed the data; MD and MH led
the writing of the manuscript; All authors contributed critically to the
drafts and gave final approval for publication.

\hypertarget{data-and-code-availability}{%
\section*{Data and Code Availability}\label{data-and-code-availability}}
\addcontentsline{toc}{section}{Data and Code Availability}

This manuscript has a supplementary \textbf{\textsf{R}} package that
contains all of the data and code used in its creation. The
supplementary \textbf{\textsf{R}} package is hosted on GitHub.
Instructions for download are available at

\url{https://github.com/michaeldumelle/DvMsp}.

If the manuscript is accepted, this repository will be archived in
Zenodo.

\hypertarget{supporting-information}{%
\section*{Supporting Information}\label{supporting-information}}
\addcontentsline{toc}{section}{Supporting Information}

In the supporting information, we provide tables of summary statistics
for all 36 simulation scenarios and all six real data scenarios.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-akaike1974new}{}%
Akaike, H., 1974. A new look at the statistical model identification.
IEEE Transactions on Automatic Control 19, 716--723.

\leavevmode\hypertarget{ref-barabesi2011sampling}{}%
Barabesi, L., Franceschi, S., 2011. Sampling properties of spatial total
estimators under tessellation stratified designs. Environmetrics 22,
271--278.

\leavevmode\hypertarget{ref-benedetti2017spatially}{}%
Benedetti, R., Piersimoni, F., 2017. A spatially balanced design with
probability function proportional to the within sample distance.
Biometrical Journal 59, 1067--1084.

\leavevmode\hypertarget{ref-benedetti2017spatiallyreview}{}%
Benedetti, R., Piersimoni, F., Postiglione, P., 2017. Spatially balanced
sampling: A review and a reappraisal. International Statistical Review
85, 439--454.

\leavevmode\hypertarget{ref-bolker2009generalized}{}%
Bolker, B.M., Brooks, M.E., Clark, C.J., Geange, S.W., Poulsen, J.R.,
Stevens, M.H.H., White, J.-S.S., 2009. Generalized linear mixed models:
A practical guide for ecology and evolution. Trends in ecology \&
evolution 24, 127--135.

\leavevmode\hypertarget{ref-breiman2001random}{}%
Breiman, L., 2001. Random forests. Machine Learning 45, 5--32.

\leavevmode\hypertarget{ref-brus1997random}{}%
Brus, D., De Gruijter, J., 1997. Random sampling or geostatistical
modelling? Choosing between design-based and model-dased sampling
strategies for soil (with discussion). Geoderma 80, 1--44.

\leavevmode\hypertarget{ref-brus2021statistical}{}%
Brus, D.J., 2021. Statistical approaches for spatial sample survey:
Persistent misconceptions and new developments. European Journal of Soil
Science 72, 686--703.

\leavevmode\hypertarget{ref-brus1993design}{}%
Brus, D.J., DeGruijter, J.J., 1993. Design-based versus model-based
estimates of spatial means: Theory and application in environmental soil
science. Environmetrics 4, 123--152.

\leavevmode\hypertarget{ref-chan2020bayesian}{}%
Chan-Golston, A.M., Banerjee, S., Handcock, M.S., 2020. Bayesian
inference for finite populations under spatial process settings.
Environmetrics 31, e2606.

\leavevmode\hypertarget{ref-chiles1999geostatistics}{}%
Chiles, J.-P., Delfiner, P., 1999. Geostatistics: Modeling Spatial
Uncertainty. John Wiley \& Sons, New York.

\leavevmode\hypertarget{ref-cicchitelli2012model}{}%
Cicchitelli, G., Montanari, G.E., 2012. Model-assisted estimation of a
spatial population mean. International Statistical Review 80, 111--126.

\leavevmode\hypertarget{ref-cooper2006sampling}{}%
Cooper, C., 2006. Sampling and variance estimation on continuous
domains. Environmetrics 17, 539--553.

\leavevmode\hypertarget{ref-cressie1993statistics}{}%
Cressie, N., 1993. Statistics for spatial data. John Wiley \& Sons.

\leavevmode\hypertarget{ref-de1990model}{}%
De Gruijter, J., Ter Braak, C., 1990. Model-free estimation from spatial
samples: A reappraisal of classical sampling theory. Mathematical
Geology 22, 407--415.

\leavevmode\hypertarget{ref-diggle2010geostatistical}{}%
Diggle, P.J., Menezes, R., Su, T.-l., 2010. Geostatistical inference
under preferential sampling. Journal of the Royal Statistical Society:
Series C (Applied Statistics) 59, 191--232.

\leavevmode\hypertarget{ref-dumelle2022spsurvey}{}%
Dumelle, M., Kincaid, T.M., Olsen, A.R., Weber, M.H., 2022. Spsurvey:
Spatial sampling design and analysis.

\leavevmode\hypertarget{ref-fix1989discriminatory}{}%
Fix, E., Hodges, J.L., 1989. Discriminatory analysis. Nonparametric
discrimination: Consistency properties. International Statistical
Review/Revue Internationale de Statistique 57, 238--247.

\leavevmode\hypertarget{ref-grafstrom2012spatiallypoisson}{}%
Grafström, A., 2012. Spatially correlated poisson sampling. Journal of
Statistical Planning and Inference 142, 139--147.

\leavevmode\hypertarget{ref-grafstrom2013well}{}%
Grafström, A., Lundström, N.L., 2013. Why well spread probability
samples are balanced. Open Journal of Statistics 3, 36--41.

\leavevmode\hypertarget{ref-grafstrom2012spatially}{}%
Grafström, A., Lundström, N.L., Schelin, L., 2012. Spatially balanced
sampling through the pivotal method. Biometrics 68, 514--520.

\leavevmode\hypertarget{ref-grafstrom2018spatially}{}%
Grafström, A., Matei, A., 2018. Spatially balanced sampling of
continuous populations. Scandinavian Journal of Statistics 45, 792--805.

\leavevmode\hypertarget{ref-hansen1983evaluation}{}%
Hansen, M.H., Madow, W.G., Tepping, B.J., 1983. An evaluation of
model-dependent and probability-sampling inferences in sample surveys.
Journal of the American Statistical Association 78, 776--793.

\leavevmode\hypertarget{ref-harville1977maximum}{}%
Harville, D.A., 1977. Maximum likelihood approaches to variance
component estimation and to related problems. Journal of the American
Statistical Association 72, 320--338.

\leavevmode\hypertarget{ref-higham2021sptotal}{}%
Higham, M., Ver Hoef, J., Frank, B., Dumelle, M., 2021. Sptotal:
Predicting totals and weighted sums from spatial data.

\leavevmode\hypertarget{ref-higham2021adjusting}{}%
Higham, M., Ver Hoef, J., Madsen, L., Aderman, A., 2021. Adjusting a
finite population block kriging estimator for imperfect detection.
Environmetrics 32, e2654.

\leavevmode\hypertarget{ref-hofman2021many}{}%
Hofman, S.C., Brus, D., 2021. How many sampling points are needed to
estimate the mean nitrate-n content of agricultural fields? A
geostatistical simulation approach with uncertain variograms. Geoderma
385, 114816.

\leavevmode\hypertarget{ref-horvitz1952generalization}{}%
Horvitz, D.G., Thompson, D.J., 1952. A generalization of sampling
without replacement from a finite universe. Journal of the American
Statistical Association 47, 663--685.

\leavevmode\hypertarget{ref-lohr2009sampling}{}%
Lohr, S.L., 2009. Sampling: Design and analysis. Nelson Education.

\leavevmode\hypertarget{ref-patterson1971recovery}{}%
Patterson, H.D., Thompson, R., 1971. Recovery of inter-block information
when block sizes are unequal. Biometrika 58, 545--554.

\leavevmode\hypertarget{ref-robertson2013bas}{}%
Robertson, B., Brown, J., McDonald, T., Jaksons, P., 2013. BAS: Balanced
acceptance sampling of natural resources. Biometrics 69, 776--784.

\leavevmode\hypertarget{ref-robertson2018halton}{}%
Robertson, B., McDonald, T., Price, C., Brown, J., 2018. Halton
iterative partitioning: Spatially balanced sampling via partitioning.
Environmental and Ecological Statistics 25, 305--323.

\leavevmode\hypertarget{ref-sarndal2003model}{}%
Särndal, C.-E., Swensson, B., Wretman, J., 2003. Model assisted survey
sampling. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-schabenberger2017statistical}{}%
Schabenberger, O., Gotway, C.A., 2017. Statistical methods for spatial
data analysis. CRC press.

\leavevmode\hypertarget{ref-sen1953estimate}{}%
Sen, A.R., 1953. On the estimate of the variance in sampling with
varying probabilities. Journal of the Indian Society of Agricultural
Statistics 5, 127.

\leavevmode\hypertarget{ref-sterba2009alternative}{}%
Sterba, S.K., 2009. Alternative model-based and design-based frameworks
for inference from samples to populations: From polarization to
integration. Multivariate Behavioral Research 44, 711--740.

\leavevmode\hypertarget{ref-stevens2003variance}{}%
Stevens, D.L., Olsen, A.R., 2003. Variance estimation for spatially
balanced samples of environmental resources. Environmetrics 14,
593--610.

\leavevmode\hypertarget{ref-stevens2004spatially}{}%
Stevens, D.L., Olsen, A.R., 2004. Spatially balanced sampling of natural
resources. Journal of the American Statistical Association 99, 262--278.

\leavevmode\hypertarget{ref-USEPA2012NLA}{}%
USEPA, 2012. National lakes assessment 2012.
https://www.epa.gov/national-aquatic-resource-surveys/national-results-and-regional-highlights-national-lakes-assessment.

\leavevmode\hypertarget{ref-verhoef2002sampling}{}%
Ver Hoef, J., 2002. Sampling and geostatistics for spatial data.
Ecoscience 9, 152--161.

\leavevmode\hypertarget{ref-verhoef2008spatial}{}%
Ver Hoef, J.M., 2008. Spatial methods for plot-based sampling of
wildlife populations. Environmental and Ecological Statistics 15, 3--13.

\leavevmode\hypertarget{ref-ver2010moving}{}%
Ver Hoef, J.M., Peterson, E.E., 2010. A moving average approach for
spatial statistical models of stream networks. Journal of the American
Statistical Association 105, 6--18.

\leavevmode\hypertarget{ref-ver2013comparison}{}%
Ver Hoef, J.M., Temesgen, H., 2013. A comparison of the spatial linear
model to nearest neighbor (k-nn) methods for forestry applications. PlOS
ONE 8, e59129.

\leavevmode\hypertarget{ref-walvoort2010r}{}%
Walvoort, D.J., Brus, D., De Gruijter, J., 2010. An r package for
spatial coverage sampling and random sampling from compact geographical
strata by k-means. Computers \& geosciences 36, 1261--1267.

\leavevmode\hypertarget{ref-wang2013design}{}%
Wang, J.-F., Jiang, C.-S., Hu, M.-G., Cao, Z.-D., Guo, Y.-S., Li, L.-F.,
Liu, T.-J., Meng, B., 2013. Design-based spatial sampling: Theory and
implementation. Environmental Modelling \& Software 40, 280--288.

\leavevmode\hypertarget{ref-wolfinger1994computing}{}%
Wolfinger, R., Tobias, R., Sall, J., 1994. Computing gaussian
likelihoods and their derivatives for general linear mixed models. SIAM
Journal on Scientific Computing 15, 1294--1310.

\leavevmode\hypertarget{ref-yates1953selection}{}%
Yates, F., Grundy, P.M., 1953. Selection without replacement from within
strata with probability proportional to size. Journal of the Royal
Statistical Society: Series B (Methodological) 15, 253--261.

\end{document}